### 2019-03-18 thesis_proposal_1/Chapters/4_Objectives.tex

% Chapter Template

\chapter{Objectives} % Main chapter title

\label{Chapter:Objectives}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

%-----------------------------------
%	SUBSECTION 1
%-----------------------------------