### 2019-03-18 thesis_proposal_1/Chapters/5_TheoreticalFramework.tex

% Chapter Template

\chapter{Theoretical Framework} % Main chapter title

\label{Chapter:TheoreticalFramework}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

%-----------------------------------
%	SUBSECTION 1
%-----------------------------------