### 2019-03-18 thesis_proposal_1/main.tex
another draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Masters/Doctoral Thesis 
% LaTeX Template
% Version 2.5 (27/8/17)
%
% This template was downloaded from:
% http://www.LaTeXTemplates.com
%
% Version 2.x major modifications by:
% Vel (vel@latextemplates.com)
%
% This template is based on a template by:
% Steve Gunn (http://users.ecs.soton.ac.uk/srg/softwaretools/document/templates/)
% Sunil Patel (http://www.sunilpatel.co.uk/thesis-template/)
%
% Template license:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
% The default document font size, options: 10pt, 11pt, 12pt
11pt, 
% Two side (alternating margins) for binding by default, uncomment to switch to one side
%oneside,
% ngerman for German
english,
% Single line spacing, alternatives: onehalfspacing or doublespacing
singlespacing,
% Uncomment to enable draft mode (no pictures, no links, overfull hboxes indicated)
%draft,
% If the document is onehalfspacing or doublespacing, uncomment this to set spacing in lists to single
%nolistspacing,
% Uncomment to add the list of figures/tables/etc to the table of contents
%liststotoc,
% Uncomment to add the main table of contents to the table of contents
%toctotoc,
% Uncomment to add space between paragraphs
parskip,
% Uncomment to not load the hyperref package
%nohyperref,
% Uncomment to get a line under the header
headsepline,
% Uncomment to place the chapter title next to the number on one line
%chapterinoneline,
% Uncomment to change the layout of the declaration, abstract and acknowledgements pages to match the default layout
%consistentlayout,
% The class file specifying the document structure
]{MastersDoctoralThesis}

% Required for inputting international characters
\usepackage[utf8]{inputenc}

% Output font encoding for international characters
\usepackage[T1]{fontenc}

% Use the Palatino font by default
\usepackage{mathpazo}

% Use the bibtex backend with the authoryear citation style (which resembles APA)
\usepackage[backend=biber,style=authoryear,natbib=true]{biblatex}
% MLA, APA, or IEEE? - https://www.overleaf.com/learn/latex/Biblatex_citation_styles
%\usepackage[style=apa, backend=biber]{biblatex}

% The filename of the bibliography
\addbibresource{example.bib}

% Required to generate language-dependent quotes in the bibliography
\usepackage[autostyle=true]{csquotes}

%----------------------------------------------------------------------------------------
%	MARGIN SETTINGS
%----------------------------------------------------------------------------------------

\geometry{
	paper=a4paper,      % Change to letterpaper for US letter
	inner=2.5cm,        % Inner margin
	outer=3.8cm,        % Outer margin
	bindingoffset=.5cm, % Binding offset
	top=1.5cm,          % Top margin
	bottom=1.5cm,       % Bottom margin
	%showframe,         % Uncomment to show how the type block is set on the page
}

%----------------------------------------------------------------------------------------
%	THESIS INFORMATION
%----------------------------------------------------------------------------------------

% Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
\thesistitle{Design of Polymer Solutions for the Fabrication of Conducting Carbon Nano-wires}

% Your supervisor's name, this is used in the title page, print it elsewhere with \supname and \cosupname
\supervisor{Dr. Héctor Alán \textsc{Aguirre} Soto}
\cosupervisor{Dra. Dora Iliana \textsc{Medina} Medina}

% Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
\examiner{}

% Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
\degree{Master of Science in Nanotechnology (MNT)}

% Your name, this is used in the title page and abstract, print it elsewhere with \authorname
\author{Antonio Osamu \textsc{Katagiri} Tanaka}

% Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname
\addresses{Estado de México, Atizapan de Zaragoza}

% Your subject area, this is not currently used anywhere in the template, print it elsewhere with \subjectname
\subject{Nanotechnology}

% Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
\keywords{tec, de, monterrey}

% Your university's name and URL, this is used in the title page and abstract, print it elsewhere with \univname
\university{\href{https://tec.mx/}{Instituto Tecnonólogico y de Estudios Superiores de Monterrey}}

% Your university's campus name and URL, this is used in the title page and abstract, print it elsewhere with \campusname
\campus{\href{https://tec.mx/}{Campus Estado de México}}

% Your department's name and URL, this is used in the title page and abstract, print it elsewhere with \deptname
\department{\href{https://tec.mx/}{School of Engineering and Sciences}}

% Your research group's name and URL, this is used in the title page, print it elsewhere with \groupname
\group{\href{https://www.medinadora.com/people}{ITESM \campusname}}

% Your faculty's name and URL, this is used in the title page and abstract, print it elsewhere with \facname
\faculty{\href{https://samp.itesm.mx/Programas/VistaPrograma?clave=MNT16&modoVista=Default&idioma=ES&cols=0}{Faculty: Nanotechnology}}

\AtBeginDocument{
\hypersetup{pdftitle=\ttitle}          % Set the PDF's title to your title
\hypersetup{pdfauthor=\authorname}     % Set the PDF's author to your name
\hypersetup{pdfkeywords=\keywordnames} % Set the PDF's keywords to your keywords
}

\begin{document}

% Use roman page numbering style (i, ii, iii, iv...) for the pre-content pages
\frontmatter

% Default to the plain heading style until the thesis style is called for the body content
\pagestyle{plain}

%----------------------------------------------------------------------------------------
%	TITLE/COVER PAGE
%----------------------------------------------------------------------------------------

\begin{titlepage}
\begin{center}

\vspace*{.06\textheight}
{\scshape\LARGE \univname\par}\vspace{1.5cm} % University name
\textsc{\Large Masters Thesis Proposal}\\[0.5cm]      % Thesis type

\HRule \\[0.4cm]                            % Horizontal line
{\huge \bfseries \ttitle\par}\vspace{0.4cm} % Thesis title
\HRule \\[1.5cm]                            % Horizontal line

\begin{minipage}[t]{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\

% Author name - remove the \href bracket to remove the link
\href{https://linkedin.com/in/osamu-katagiri-84b2b940/}{\authorname}
%\authorname
\end{flushleft}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\begin{flushright} \large
\emph{Principal Advisor:} \\
% Supervisor name - remove the \href bracket to remove the link
%\href{http://www.doramedina.com}{\supname}
\supname

\bigskip

\emph{Co-advisor and\\Director of Program:} \\
% Supervisor name - remove the \href bracket to remove the link
\href{https://www.medinadora.com/}{\cosupname}
%\cosupname
\end{flushright}
\end{minipage}\\[3cm]

\vfill

\large \textit{A thesis submitted in fulfillment of the requirements\\ for the degree of
% University requirement text
\degreename}\\[0.3cm]
\textit{in}\\[0.4cm]
% Research group name and department name
\groupname\\\deptname\\[2cm]

\vfill

% Address and Date
{\addressname \text{, } \large \today}\\[4cm]
% University/department logo - uncomment to place it
%\includegraphics{./Figures/uniLogo.png}

\vfill
\end{center}
\end{titlepage}

%----------------------------------------------------------------------------------------
%	COMMITTEE'S PAGE
%----------------------------------------------------------------------------------------

\begin{committee}
\addchaptertocentry{\supervisingcommittee} % Add the the supervising committee to the table of contents
The committee members, hereby, recommend that the proposal by \authorname to be accepted to develop the thesis project as a partial requirement for the degree of \degreename.

\vfill

\begin{flushright} \large

\bigskip
\bigskip
\bigskip
\bigskip
\bigskip

\noindent \rule[0.5em]{15em}{0.5pt}\\ % This prints a line for the signature
%\href{http://www.alanaguirre.com}{\supname}
\supname\\
Tecnológico de Monterrey\\
\emph{Principal Advisor}

\bigskip
\bigskip
\bigskip
\bigskip
\bigskip

\noindent \rule[0.5em]{15em}{0.5pt}\\ % This prints a line for the signature
%\href{http://www.alanaguirre.com}{\supname}
1st committee member name\\
1st committee member institution\\
\emph{Committee Member}

\bigskip
\bigskip
\bigskip
\bigskip
\bigskip

\noindent \rule[0.5em]{15em}{0.5pt}\\ % This prints a line for the signature
%\href{http://www.alanaguirre.com}{\supname}
2nd committee member name\\
2nd committee member institution\\
\emph{Committee Member}

\end{flushright}

\vfill

\begin{center}

\bigskip
\bigskip
\bigskip
\bigskip
\bigskip

\noindent \rule[0.5em]{15em}{0.5pt}\\ % This prints a line for the signature
\cosupname \\
\emph{Director of Program in Nanotechnology}\\
School of Engineering and Sciences\\
%\href{http://www.doramedina.com}{\supname}

\bigskip
\bigskip
{\addressname \text{, } \large \today}\\[4cm]

\end{center}
\end{committee}

\cleardoublepage

%----------------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES PAGES
%----------------------------------------------------------------------------------------

\tableofcontents % Prints the main table of contents

%\listoffigures % Prints the list of figures

%\listoftables % Prints the list of tables

%----------------------------------------------------------------------------------------
%	ABSTRACT PAGE
%----------------------------------------------------------------------------------------

\begin{abstract}
\addchaptertocentry{\abstractname} % Add the abstract to the table of contents

Carbon nano-wires are versatile materials composed of carbon chains with a wide range of applications due to the  //

Over the last couple of decades, Carbon Nanowires (CNWs) have found important applications in the field of rapid label-free biosensing platforms for their excellent electrical, thermal and chemical properties. Despite the growing interest in incorporating these materials into sensitive analytical devices, there has been limited research efforts that explore the realistic incorporation of these structures into broader microsystems. The present dissertation explores two Advanced Additive Manufacturing (AdAM) techniques for the fabrication of CNWs with promising characteristics for sensing applications. The first technique, Electro-Mechanical Spinning (EMS) is a relatively new method for the production of CNWs with highly-oriented patterning capabilities, and is assessed in terms of the minimum feature sizes that it can attain reproducibly. The second method, Multiple Photon Polymerization (MPP) is a mature laser manufacturing technique that offers high repeatability, yet it has remained largely unexplored in the field of carbon materials. This work presents a study of how MPP can be used to create CNWs in terms of micro-fabrication parameters like laser power and scanning velocity. A thorough review study of the surface properties required for carbon materials to become eligible candidates in biosensing is also included in this dissertation. Lastly, an application of CNWs is presented, in which a Joule Heating technique is used to induce controlled breaking of the wires to produce nano-scale separated carbon electrodes, which hold potential for micro/nano particle manipulation in microfluidics. //

Carbon fibers are versatile materials composed of carbon chains with a wide range of applications. Carbon nanowires have been fabricated with a photoresist, but little is known about polymers that can produce more conductive carbon nanowires after pyrolysis. FFES or Far field electrospinning is a inexpensive process to produce polymer fibers, but with low accuracy due to electric instabilities. Near field electrospinning (NFES) is similar to FFES, but with a reduced distance between the fiber source and the collector and higher precision. Various polymer solutions have being tested and measured through NFSE and photopolymerization processes; it was found that it is not possible to predict the behaviour of the electrospinning process, so additional properties are to be considered to achieve a stable process. The properties of the polymer solution can be amended to obtain a electrospun-able, photopolymerizable and pyrolyzable/graphitizable fibers. Analyse the rheology of different polymer solutions to determine if they can be easily electrospun at low voltages and then fabricate nanowires with them. Design polymer solutions that can be electrospun by NFES, photopolymerized, and then pyrolyzed into conducting carbon nanowires. The use of the newly designed polymer solution within the NFES-Photopolymerization-Pyrolysis process can be amended to achieve mass scale manufacturing of carbon nanowires in a cheap, continuous, simple and reproducible manner. The new/tailored polymer solution can be used to synthesize carbon nanowires with high conductive properties and reduced thickness. \\

\noindent
\textbf{keywords:} \keywordnames
\end{abstract}

%----------------------------------------------------------------------------------------
%	ABBREVIATIONS
%----------------------------------------------------------------------------------------

%\begin{abbreviations}{ll} % Include a list of abbreviations (a table of two columns)

%\textbf{ITESM} & \textbf{I}nstituto \textbf{T}ecnonólogico y de \textbf{E}studios \textbf{S}uperiores de \textbf{M}onterrey\\

%\end{abbreviations}

%----------------------------------------------------------------------------------------
%	PHYSICAL CONSTANTS/OTHER DEFINITIONS
%----------------------------------------------------------------------------------------

%\begin{constants}{lr@{${}={}$}l} % The list of physical constants is a three column table

% The \SI{}{} command is provided by the siunitx package, see its documentation for instructions on how to use it

%Speed of Light & $c_{0}$ & \SI{2.99792458e8}{\meter\per\second} (exact)\\
%Constant Name & $Symbol$ & $Constant Value$ with units\\

%\end{constants}

%----------------------------------------------------------------------------------------
%	SYMBOLS
%----------------------------------------------------------------------------------------

%\begin{symbols}{lll} % Include a list of Symbols (a three column table)

%$a$ & distance & \si{\meter} \\
%$P$ & power & \si{\watt} (\si{\joule\per\second}) \\
%Symbol & Name & Unit \\

%\addlinespace % Gap to separate the Roman symbols from the Greek

%$\omega$ & angular frequency & \si{\radian} \\

%\end{symbols}

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - CHAPTERS
%----------------------------------------------------------------------------------------

\mainmatter % Begin numeric (1,2,3...) page numbering

\pagestyle{thesis} % Return the page headers back to the "thesis" style

% Include the chapters of the thesis as separate files from the Chapters folder
% Uncomment the lines as you write the chapters

\include{Chapters/1_Introduction}
\include{Chapters/2_ProblmemDefinitionAndMotivation} 
\include{Chapters/3_HypothesisAndResearchQuestions}
\include{Chapters/4_Objectives} 
\include{Chapters/5_TheoreticalFramework}
\include{Chapters/6_Methodology}
\include{Chapters/7_WorkPlan}

%\include{Chapters/__IntroToLaTeX}

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - APPENDICES
%----------------------------------------------------------------------------------------

\appendix % Cue to tell LaTeX that the following "chapters" are Appendices

% Include the appendices of the thesis as separate files from the Appendices folder
% Uncomment the lines as you write the Appendices

%\include{Appendices/AppendixA}
%\include{Appendices/AppendixB}
%\include{Appendices/AppendixC}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\printbibliography[title={References}]

%----------------------------------------------------------------------------------------

\end{document}