### 2019-03-19 thesis_proposal_1/Chapters/3_HypothesisAndResearchQuestions.tex

% Chapter Template

\chapter{Hypothesis and Research Questions} % Main chapter title

\label{Chapter:HypothesisandResearchQuestions}

\section{Research Hypothesis}

The rheological properties of polymer solutions along with synthesis parameters (stage velocity, voltage, dispense rate) can be amended to obtain a low voltage electrospun-able, photopolymerizable and graphitizable fibers to synthesize carbon nano-wires with specified dimensions (diameter and length).

\section{Research Questions}

\begin{itemize}
	\item{
	Are there any evidence of carbon nano-wire fabrication though electrospun-able and pyrozable polymer solutions?
	}
	\item{
	What are the process parameters to consider/control for the fabrication processes of carbon nano-wires? 
	}
	\item{
	What rheological properties are to be controlled/tested to deliver a electrospun-able and pyrozable polymer solution?	
	}
	\item{
	Are there any efforts employed to the design of polymer solutions that can be electrospun, photopolymerized, and pyrolyzed into conducting carbon nanowires?
	}
	\item{
	Are the fabrication parameters defined in \cite{Cardenas2017} optimal for the synthesis of carbon nano-wires through near-field electromechanical spinning?	
	}
\end{itemize}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

%-----------------------------------
%	SUBSECTION 1
%-----------------------------------