### 2019-03-19 thesis_proposal_1/Chapters/4_Objectives.tex

% Chapter Template

\chapter{Objectives} % Main chapter title

\label{Chapter:Objectives}

\section{General objective}
Exercise the practice and feasibility of a new synthesis process to achieve mass scale manufacturing of carbon nano-wires in an inexpensive, continuous, simple and reproducible manner.

\section{Particular objective}
Design polymer solutions that can be electrospun by NFES, photopolymerized, and then pyrolyzed; through rheological analyses to determine if they can be easily employed for conducting carbon nano-wire synthesis.

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

%-----------------------------------
%	SUBSECTION 1
%-----------------------------------