### 2019-03-19 thesis_proposal_1/main.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Masters/Doctoral Thesis 
% LaTeX Template
% Version 2.5 (27/8/17)
%
% This template was downloaded from:
% http://www.LaTeXTemplates.com
%
% Version 2.x major modifications by:
% Vel (vel@latextemplates.com)
%
% This template is based on a template by:
% Steve Gunn (http://users.ecs.soton.ac.uk/srg/softwaretools/document/templates/)
% Sunil Patel (http://www.sunilpatel.co.uk/thesis-template/)
%
% Template license:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
% The default document font size, options: 10pt, 11pt, 12pt
11pt, 
% Two side (alternating margins) for binding by default, uncomment to switch to one side
%oneside,
% ngerman for German
english,
% Single line spacing, alternatives: singlespacing, onehalfspacing or doublespacing
onehalfspacing,
% Uncomment to enable draft mode (no pictures, no links, overfull hboxes indicated)
%draft,
% If the document is nolistspacing, onehalfspacing or doublespacing, uncomment this to set spacing in lists to single
onehalfspacing,
% Uncomment to add the list of figures/tables/etc to the table of contents
%liststotoc,
% Uncomment to add the main table of contents to the table of contents
%toctotoc,
% Uncomment to add space between paragraphs
parskip,
% Uncomment to not load the hyperref package
%nohyperref,
% Uncomment to get a line under the header
headsepline,
% Uncomment to place the chapter title next to the number on one line
chapterinoneline,
% Uncomment to change the layout of the declaration, abstract and acknowledgements pages to match the default layout
%consistentlayout,
% The class file specifying the document structure
]{MastersDoctoralThesis}

% Required for inputting international characters
\usepackage[utf8]{inputenc}

% Output font encoding for international characters
\usepackage[T1]{fontenc}

% Use the Palatino font by default
\usepackage{mathpazo}

% Use tu justify paragraphs
\usepackage{ragged2e}

% Use the bibtex backend with the authoryear citation style (which resembles APA)
\usepackage[backend=biber,natbib=true]{biblatex}
% MLA, APA, or IEEE? - https://www.overleaf.com/learn/latex/Biblatex_citation_styles
%\usepackage[style=apa, backend=biber]{biblatex}

% The filename of the bibliography
\addbibresource{mainBiblio.bib}

% Required to generate language-dependent quotes in the bibliography
\usepackage[autostyle=true]{csquotes}

%----------------------------------------------------------------------------------------
%	MARGIN SETTINGS
%----------------------------------------------------------------------------------------

\geometry{
	paper=a4paper,      % Change to letterpaper for US letter
	inner=2.5cm,        % Inner margin
	outer=3.8cm,        % Outer margin
	bindingoffset=.5cm, % Binding offset
	top=1.5cm,          % Top margin
	bottom=1.5cm,       % Bottom margin
	%showframe,         % Uncomment to show how the type block is set on the page
}

%----------------------------------------------------------------------------------------
%	THESIS INFORMATION
%----------------------------------------------------------------------------------------

% Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
\thesistitle{Design of Polymer Solutions for the Fabrication of Conducting Carbon Nano-wires}

% Your supervisor's name, this is used in the title page, print it elsewhere with \supname and \cosupname
\supervisor{Dr. Héctor Alán \textsc{Aguirre} Soto}
\cosupervisor{Dra. Dora Iliana \textsc{Medina} Medina}

% Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
\examiner{}

% Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
\degree{Master of Science in Nanotechnology (MNT)}

% Your name, this is used in the title page and abstract, print it elsewhere with \authorname
\author{Antonio Osamu \textsc{Katagiri} Tanaka}

% Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname
\addresses{Estado de México, Atizapan de Zaragoza}

% Your subject area, this is not currently used anywhere in the template, print it elsewhere with \subjectname
\subject{Nanotechnology}

% Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
\keywords{nanotechnology, carbon, nano-wires, electrospinning, NFES}

% Your university's name and URL, this is used in the title page and abstract, print it elsewhere with \univname
\university{\href{https://tec.mx/}{Instituto Tecnonólogico y de Estudios Superiores de Monterrey}}

% Your university's campus name and URL, this is used in the title page and abstract, print it elsewhere with \campusname
\campus{\href{https://tec.mx/}{Campus Estado de México}}

% Your department's name and URL, this is used in the title page and abstract, print it elsewhere with \deptname
\department{\href{https://tec.mx/}{School of Engineering and Sciences}}

% Your research group's name and URL, this is used in the title page, print it elsewhere with \groupname
\group{\href{https://www.medinadora.com/people}{ITESM \campusname}}

% Your faculty's name and URL, this is used in the title page and abstract, print it elsewhere with \facname
\faculty{\href{https://samp.itesm.mx/Programas/VistaPrograma?clave=MNT16&modoVista=Default&idioma=ES&cols=0}{Faculty: Nanotechnology}}

\AtBeginDocument{
\hypersetup{pdftitle=\ttitle}          % Set the PDF's title to your title
\hypersetup{pdfauthor=\authorname}     % Set the PDF's author to your name
\hypersetup{pdfkeywords=\keywordnames} % Set the PDF's keywords to your keywords
}

\begin{document}

% Use roman page numbering style (i, ii, iii, iv...) for the pre-content pages
\frontmatter

% Default to the plain heading style until the thesis style is called for the body content
\pagestyle{plain}

%----------------------------------------------------------------------------------------
%	TITLE/COVER PAGE
%----------------------------------------------------------------------------------------

\begin{titlepage}
\begin{center}

\vspace*{.06\textheight}
{\scshape\LARGE \univname\par} % University name

\begin{center}
\includegraphics[width=0.15\textwidth]{./Figures/uniLogo.png}
\end{center}

\vspace{0.5cm}
\textsc{\Large Masters Thesis Proposal}\\[0.5cm]      % Thesis type

\HRule \\[0.5cm]                            % Horizontal line
{\huge \bfseries \ttitle\par}\vspace{0.25cm} % Thesis title
\HRule \\[0.5cm]                            % Horizontal line

\begin{minipage}[t]{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\

% Author name - remove the \href bracket to remove the link
\href{https://linkedin.com/in/osamu-katagiri-84b2b940/}{\authorname}
%\authorname
\end{flushleft}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\begin{flushright} \large
\emph{Principal Advisor:} \\
% Supervisor name - remove the \href bracket to remove the link
%\href{http://www.doramedina.com}{\supname}
\supname

\bigskip

\emph{Co-advisor and\\Director of Program:} \\
% Supervisor name - remove the \href bracket to remove the link
\href{https://www.medinadora.com/}{\cosupname}
%\cosupname
\end{flushright}
\end{minipage}\\[0.5cm]

\vfill

\large \textit{A thesis proposal submitted in fulfillment of the requirements\\ for the degree of
% University requirement text
\degreename}\\[0.25cm]
\textit{in}\\[0.25cm]
% Research group name and department name
\groupname\\\deptname\\[1cm]

\vfill

% Address and Date
{\addressname \text{, } \large \today}\\[4cm]
% University/department logo - uncomment to place it
%\includegraphics{./Figures/uniLogo.png}

\vfill
\end{center}
\end{titlepage}

%----------------------------------------------------------------------------------------
%	COMMITTEE'S PAGE
%----------------------------------------------------------------------------------------

\begin{committee}
\addchaptertocentry{\supervisingcommittee} % Add the the supervising committee to the table of contents
The committee members, hereby, recommend that the proposal by \authorname to be accepted to develop the thesis project as a partial requirement for the degree of \degreename.

\vfill

\begin{flushright} \large

\bigskip
\bigskip
\bigskip
\bigskip

\noindent \rule[0.5em]{15em}{0.5pt}\\ % This prints a line for the signature
%\href{http://www.alanaguirre.com}{\supname}
\supname\\
Tecnológico de Monterrey\\
\emph{Principal Advisor}

\bigskip
\bigskip
\bigskip
\bigskip

\noindent \rule[0.5em]{15em}{0.5pt}\\ % This prints a line for the signature
%\href{http://www.alanaguirre.com}{\supname}
1st committee member name\\
1st committee member institution\\
\emph{Committee Member}

\bigskip
\bigskip
\bigskip
\bigskip

\noindent \rule[0.5em]{15em}{0.5pt}\\ % This prints a line for the signature
%\href{http://www.alanaguirre.com}{\supname}
2nd committee member name\\
2nd committee member institution\\
\emph{Committee Member}

\end{flushright}

\vfill

\begin{center}

\bigskip
\bigskip
\bigskip
\bigskip

\noindent \rule[0.5em]{15em}{0.5pt}\\ % This prints a line for the signature
\cosupname \\
\emph{Director of Program in Nanotechnology}\\
School of Engineering and Sciences\\
%\href{http://www.doramedina.com}{\supname}

\bigskip
\bigskip
{\addressname \text{, } \large \today}\\[4cm]

\end{center}
\end{committee}

\cleardoublepage

%----------------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES PAGES
%----------------------------------------------------------------------------------------

\tableofcontents % Prints the main table of contents

%\listoffigures % Prints the list of figures

%\listoftables % Prints the list of tables

%----------------------------------------------------------------------------------------
%	ABSTRACT PAGE
%----------------------------------------------------------------------------------------

\begin{abstract}
\addchaptertocentry{\abstractname} % Add the abstract to the table of contents

Carbon nano-wires are versatile materials composed of carbon chains with a wide range of applications due to their matchless properties in electrical matters. Regardless of the high interest in the implementation of carbon nano-wires in several applications and devices, no feasible processes have been developed to synthesize carbon nano-wires. Carbon nano-wires have been fabricated with the use of a photoresist, but little is known about polymers that can produce more conductive carbon nano-wires after pyrolysis. Various polymer solutions have being tested and measured through near field electrospinning (NFES) and photopolymerization processes; it was found that it is not possible to predict the behaviour of the electrospinning process, so additional properties are to be considered to achieve a stable manufacturing process. The thesis proposal is to analyse the rheology of different polymer solutions to determine if they can be easily electrospun at low voltages and then fabricate nanowires with them. The research process will include the design of polymer solutions that can be electrospun, photopolymerized, and then pyrolyzed into conducting carbon nanowires. The research is intended to engineer a newly designed polymer solution to achieve mass scale manufacturing of carbon nano-wires in a cheap, continuous, simple and reproducible manner. \\

\noindent
\textbf{keywords:} \keywordnames
\end{abstract}

%----------------------------------------------------------------------------------------
%	ABBREVIATIONS
%----------------------------------------------------------------------------------------

%\begin{abbreviations}{ll} % Include a list of abbreviations (a table of two columns)

%\textbf{ITESM} & \textbf{I}nstituto \textbf{T}ecnonólogico y de \textbf{E}studios \textbf{S}uperiores de \textbf{M}onterrey\\

%\end{abbreviations}

%----------------------------------------------------------------------------------------
%	PHYSICAL CONSTANTS/OTHER DEFINITIONS
%----------------------------------------------------------------------------------------

%\begin{constants}{lr@{${}={}$}l} % The list of physical constants is a three column table

% The \SI{}{} command is provided by the siunitx package, see its documentation for instructions on how to use it

%Speed of Light & $c_{0}$ & \SI{2.99792458e8}{\meter\per\second} (exact)\\
%Constant Name & $Symbol$ & $Constant Value$ with units\\

%\end{constants}

%----------------------------------------------------------------------------------------
%	SYMBOLS
%----------------------------------------------------------------------------------------

%\begin{symbols}{lll} % Include a list of Symbols (a three column table)

%$a$ & distance & \si{\meter} \\
%$P$ & power & \si{\watt} (\si{\joule\per\second}) \\
%Symbol & Name & Unit \\

%\addlinespace % Gap to separate the Roman symbols from the Greek

%$\omega$ & angular frequency & \si{\radian} \\

%\end{symbols}

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - CHAPTERS
%----------------------------------------------------------------------------------------

\mainmatter % Begin numeric (1,2,3...) page numbering

\pagestyle{thesis} % Return the page headers back to the "thesis" style

% Include the chapters of the thesis as separate files from the Chapters folder
% Uncomment the lines as you write the chapters

\setlength{\parskip}{1.5em} % Paragraph spacing
%\RaggedRight
\include{Chapters/1_Introduction}
\include{Chapters/2_ProblmemDefinitionAndMotivation} 
\include{Chapters/3_HypothesisAndResearchQuestions}
\include{Chapters/4_Objectives} 
\include{Chapters/5_TheoreticalFramework}
\include{Chapters/6_Methodology}
\include{Chapters/7_WorkPlan}

%\include{Chapters/__IntroToLaTeX}

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - APPENDICES
%----------------------------------------------------------------------------------------

\appendix % Cue to tell LaTeX that the following "chapters" are Appendices

% Include the appendices of the thesis as separate files from the Appendices folder
% Uncomment the lines as you write the Appendices

%\include{Appendices/AppendixA}
%\include{Appendices/AppendixB}
%\include{Appendices/AppendixC}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\printbibliography[title={References}]

%----------------------------------------------------------------------------------------

\end{document}