### 2019-03-19 thesis_proposal_1/Chapters/2_ProblmemDefinitionAndMotivation.tex
proposal in preogress

% Chapter Template

\chapter{Problem Definition and Motivation} % Main chapter title

\label{Chapter:ProblemDefinitionandMotivation}

% Describe the current situation of this problem.
Carbon nanowires have been fabricated with a photoresist by multiple-photon polymerization techniques. However little is known about polymers that can produce conductive carbon nanowires after pyrolysis.

% identifying the obstacles and the possible applications of your research.

% background of the problem (e.g. origins, previous attempts, etc.).

% State the problem in detail mentioning all relative aspects, variables, relationships,

% stressing on the importance of finding a solution to the problem.

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

%-----------------------------------
%	SUBSECTION 1
%-----------------------------------