### 2019-04-02 thesis_proposal_LaTeX/Chapters/4_Objectives.tex

% Chapter Template

\chapter{Objectives} % Main chapter title

\label{Chapter:Objectives}

\subsubsection*{\color{mygray}[Chapter ready for review]}
% Here you are meant to stipulate, in a broadly and specific way, what you want to achieve regarding the target problem. You may include the scope of your proposed research or any other element that you consider might set the boundaries of your work.
% This section can also present the Solution Model which is the one you propose to solve the problem. Obviously, this model has to be consistent with the objectives that have already set up. This model is the one that guides the research work and that, eventually, leads to the PERSONAL CONTRIBUTION.

\section{General objective}
Study the practice and feasibility of a new synthesis process to achieve mass scale manufacturing of carbon nano-wires in an inexpensive, continuous, simple and reproducible manner; by the integration of mechano-electrospinning technique.

\section{Particular objectives}

\begin{itemize}
	\item{
	Design polymer solutions that can be electrospun by NFES, photopolymerized, and then pyrolyzed.
    }
    \item{
    Through rheological analyses, determine if polymer solutions can be easily employed for conducting carbon nano-wire synthesis.
    }
    \item{
    Determine and control the polymer solution rheological properties along with the process parameters of carbon nano-wire synthesis.
    }
\end{itemize}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

%-----------------------------------
%	SUBSECTION 1
%-----------------------------------