### 2019-04-05 thesis_proposal_LaTeX/Chapters/6_StateoftheArt.tex
related work chapter

% Chapter Template

\chapter{Related Work} % Main chapter title

\label{Chapter:RelatedWork}

%\subsubsection*{\color{mygray}[Chapter under work]}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Role of rheological properties in near field electrospun fibers morphology \cite{Flores2017}}

\section{Advanced Manufacturing Techniques for the Fabrication and Surface Modification of Carbon Nanowires \cite{Cardenas2017}}

%-----------------------------------
%	SUBSECTION 1
%-----------------------------------