### 2019-04-05 thesis_proposal_LaTeX/Chapters/6_StateoftheArt.tex
more related work

% Chapter Template

\chapter{Related Work} % Main chapter title

\label{Chapter:RelatedWork}

%\subsubsection*{\color{mygray}[Chapter under work]}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Role of rheological properties in near field electrospun fibers morphology \cite{Flores2017}}

Flores studied SU-8 2002 polymer solutions in cyclopentanone with poly(ethylene) oxide (PEO) and tetrabutylammonium tetrafluoroborate (TBATFB). For that purpose, several samples were prepared with the adequate amounts of PEO and TBATFB with 5 $m l$ of SU-8 2002 and stirred at 160 $r p m$ for one hour at 75 $^\circ C$ in the absence of light. A 5 $m l$ syringe was used to extract the solution from the preparation vial. Finally the syringe was placed upside down for 24 hours to get rid of any bubbles within the solution. Table \ref{tbl:FloresSamples} lists the set of samples that were prepared as values in $w t \%$, dissolved in SU-8 2002.

\begin{table}[th]
\caption{Set of prepared samples}
\begin{center}
\begin{tabular}{ c c c c c c c } 
\hline
\text{} & \multicolumn{2}{c}{Serie A} & \multicolumn{2}{c}{Serie B} & \multicolumn{2}{c}{Serie C} \\
\hline
Sample & PEO & TBATFB & PEO & TBATFB & PEO & TBATFB \\
\hline
1 & 0.00 & 0.00 & 0.00 & 0.50 & 0.50 & 0.00 \\
2 & 0.25 & 0.25 & 0.25 & 0.50 & 0.50 & 0.25 \\
3 & 0.50 & 0.50 & 0.50 & 0.50 & 0.50 & 0.50 \\
4 & 0.75 & 0.75 & 0.75 & 0.50 & 0.50 & 0.75 \\
5 & 1.00 & 1.00 & 1.00 & 0.50 & 0.50 & 1.00 \\
\hline
\label{tbl:FloresSamples}
\end{tabular}
\end{center}
\end{table}

All the samples were executed in a rheometer (Physica MCR 301, Anton Paar) with a cone-and-plane geometry of diameter 24.979 $m m$, angle 4.014$^\circ$ and truncation of 249 $\mu m$. The measurements were performed at a temperature of 25 $\pm$ 0.1 $^\circ C$. For amplitude sweep measurements, the angular frequency was settled at 10 $s^{-1}$, and the percentage amplitude gamma $\% \gamma$, was varied from 0.1 to 1000\%. In flow curve tests, shear rate was applied from 0.1 to 100 $s^{-1}$. For frequency sweeps, the percentage of amplitude gamma, $\% \gamma$, was varied from 0.1 to 100 $s^{-1}$. During the measurements, the rheometer sample area was saturated with cyclopentanone to avoid solvent evaporation.

\subsection{Rheological Characterization : \textbf{Amplitude Sweep}}
The "Serie A`` result measurements indicate that the storage modulus $G'$ is smaller than the loss modulus $G''$. Both moduli have a parabolic behaviour. At low deformation the values of the moduli increase until they become stable at between 1 and 10 $\% \gamma$. After the stabilization, both modulus start to decrease. The increase of PEO and TBATFB concentration $G'$ and $G''$ also increase. Figure \ref{fig:SerieAampSweep} shows the amplitude sweep for the samples of "Serie A``.

\begin{figure}[th]
\centering
\includegraphics[width=0.75\textwidth]{./Figures/SerieAampSweep.png}
\decoRule
\caption[Serie A - Amplitude Sweep]{Serie A - Amplitude Sweep}
\label{fig:SerieAampSweep}
\end{figure}

The results within Figure \ref{fig:SerieAampSweep} showed that no clear linear viscoelastic region is present. The material has influenced by small deformations, hence it is very sensible to external forces. No yield point was encountered as the moduli separate from each other with the increase of deformation.

Figure \ref{fig:SerieBampSweep} illustrates the results obtained from the amplitude sweeps of Serie B samples. As shown in the figure, the concentration of 0 $w t \%$ shows a constant viscous modulus at 0.2 $Pa$. the 0.25 $w t \%$ concentration sample presented a similar behaviour to the 0 $w t \%$ sample but with a constant value of $G'$ at 2 $Pa$. The other concentrations show a similar parabolic behaviour as the ones in Serie A.

\begin{figure}[th]
\centering
\includegraphics[width=0.75\textwidth]{./Figures/SerieBampSweep.png}
\decoRule
\caption[Serie B - Amplitude Sweep]{Serie B - Amplitude Sweep}
\label{fig:SerieBampSweep}
\end{figure}

\subsection{Rheological Characterization : \textbf{Flow Curve}}

\subsection{Rheological Characterization : \textbf{Frequency Sweep}}

\clearpage

\section{Advanced Manufacturing Techniques for the Fabrication and Surface Modification of Carbon Nanowires \cite{Cardenas2017}}

%-----------------------------------
%	SUBSECTION 1
%-----------------------------------