### 2019-06-20 NFES_Review/article.tex
update

%%
%% Copyright 2007, 2008, 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
\documentclass[3p,,preprint,12pt]{elsarticle}
\makeatletter\if@twocolumn\PassOptionsToPackage{switch}{lineno}\else\fi\makeatother

\usepackage{tabulary,xcolor}
\usepackage{amsfonts,amsmath,amssymb}
\usepackage[T1]{fontenc}
\makeatletter
\let\save@ps@pprintTitle\ps@pprintTitle
\def\ps@pprintTitle{\save@ps@pprintTitle\gdef\@oddfoot{\footnotesize\itshape \null\hfill\today}}
\def\hlinewd#1{%
  \noalign{\ifnum0=`}\fi\hrule \@height #1%
  \futurelet\reserved@a\@xhline}
\def\tbltoprule{\hlinewd{.8pt}\\[-12pt]}
\def\tblbottomrule{\noalign{\vspace*{6pt}}\hline\noalign{\vspace*{2pt}}}
\def\tblmidrule{\noalign{\vspace*{6pt}}\hline\noalign{\vspace*{2pt}}}
\AtBeginDocument{\ifNAT@numbers \biboptions{sort&compress}\fi}
\makeatother

\usepackage{ifluatex}
\ifluatex
\usepackage{fontspec}
\defaultfontfeatures{Ligatures=TeX}
\usepackage[]{unicode-math}
\unimathsetup{math-style=TeX}
\else 
\usepackage[utf8]{inputenc}
\fi 
\ifluatex\else\usepackage{stmaryrd}\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Following additional macros are required to function some 
% functions which are not available in the class used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{url,multirow,morefloats,floatflt,cancel,tfrupee}
\makeatletter

\AtBeginDocument{\@ifpackageloaded{textcomp}{}{\usepackage{textcomp}}}
\makeatother
\usepackage{colortbl}
\usepackage{xcolor}
\usepackage{pifont}
\usepackage[nointegrals]{wasysym}
\urlstyle{rm}
\makeatletter

%%%For Table column width calculation.
\def\mcWidth#1{\csname TY@F#1\endcsname+\tabcolsep}

%%Hacking center and right align for table
\def\cAlignHack{\rightskip\@flushglue\leftskip\@flushglue\parindent\z@\parfillskip\z@skip}
\def\rAlignHack{\rightskip\z@skip\leftskip\@flushglue \parindent\z@\parfillskip\z@skip}

%Etal definition in references
\@ifundefined{etal}{\def\etal{\textit{et~al}}}{}

%\if@twocolumn\usepackage{dblfloatfix}\fi
\usepackage{ifxetex}
\ifxetex\else\if@twocolumn\@ifpackageloaded{stfloats}{}{\usepackage{dblfloatfix}}\fi\fi

\AtBeginDocument{
\expandafter\ifx\csname eqalign\endcsname\relax
\def\eqalign#1{\null\vcenter{\def\\{\cr}\openup\jot\m@th
  \ialign{\strut$\displaystyle{##}$\hfil&$\displaystyle{{}##}$\hfil
      \crcr#1\crcr}}\,}
\fi
}

%For fixing hardfail when unicode letters appear inside table with endfloat
\AtBeginDocument{%
  \@ifpackageloaded{endfloat}%
   {\renewcommand\efloat@iwrite[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}}{\newif\ifefloat@tables}%
}%

\def\BreakURLText#1{\@tfor\brk@tempa:=#1\do{\brk@tempa\hskip0pt}}
\let\lt=<
\let\gt=>
\def\processVert{\ifmmode|\else\textbar\fi}
\let\processvert\processVert

\@ifundefined{subparagraph}{
\def\subparagraph{\@startsection{paragraph}{5}{2\parindent}{0ex plus 0.1ex minus 0.1ex}%
{0ex}{\normalfont\small\itshape}}%
}{}

% These are now gobbled, so won't appear in the PDF.
\newcommand\role[1]{\unskip}
\newcommand\aucollab[1]{\unskip}

\@ifundefined{tsGraphicsScaleX}{\gdef\tsGraphicsScaleX{1}}{}
\@ifundefined{tsGraphicsScaleY}{\gdef\tsGraphicsScaleY{.9}}{}
% To automatically resize figures to fit inside the text area
\def\checkGraphicsWidth{\ifdim\Gin@nat@width>\linewidth
	\tsGraphicsScaleX\linewidth\else\Gin@nat@width\fi}

\def\checkGraphicsHeight{\ifdim\Gin@nat@height>.9\textheight
	\tsGraphicsScaleY\textheight\else\Gin@nat@height\fi}

\def\fixFloatSize#1{}%\@ifundefined{processdelayedfloats}{\setbox0=\hbox{\includegraphics{#1}}\ifnum\wd0<\columnwidth\relax\renewenvironment{figure*}{\begin{figure}}{\end{figure}}\fi}{}}
\let\ts@includegraphics\includegraphics

\def\inlinegraphic[#1]#2{{\edef\@tempa{#1}\edef\baseline@shift{\ifx\@tempa\@empty0\else#1\fi}\edef\tempZ{\the\numexpr(\numexpr(\baseline@shift*\f@size/100))}\protect\raisebox{\tempZ pt}{\ts@includegraphics{#2}}}}

%\renewcommand{\includegraphics}[1]{\ts@includegraphics[width=\checkGraphicsWidth]{#1}}
\AtBeginDocument{\def\includegraphics{\@ifnextchar[{\ts@includegraphics}{\ts@includegraphics[width=\checkGraphicsWidth,height=\checkGraphicsHeight,keepaspectratio]}}}

\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}

\def\URL#1#2{\@ifundefined{href}{#2}{\href{#1}{#2}}}

%%For url break
\def\UrlOrds{\do\*\do\-\do\~\do\'\do\"\do\-}%
\g@addto@macro{\UrlBreaks}{\UrlOrds}

\edef\fntEncoding{\f@encoding}
\def\EUoneEnc{EU1}
\makeatother
\def\floatpagefraction{0.8} 
\def\dblfloatpagefraction{0.8}
\def\style#1#2{#2}
\def\xxxguillemotleft{\fontencoding{T1}\selectfont\guillemotleft}
\def\xxxguillemotright{\fontencoding{T1}\selectfont\guillemotright}

\newif\ifmultipleabstract\multipleabstractfalse%
\newenvironment{typesetAbstractGroup}{}{}%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\emergencystretch 20pt \tolerance = 1500 \def\floatpagefraction{0.8}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Feature enabled:
%toc: yes
%pagenum: yes
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\makeatletter
\def\ps@pprintTitle{\save@ps@pprintTitle\gdef\@oddfoot{\footnotesize\hspace*{.5\textwidth}\thepage\itshape \null\hfill\today}}
\makeatother
          \makeatletter\@ifundefined{tableofcontents}{\usepackage{typeset-custom-toc}}{}\makeatother
\usepackage{etoolbox}

\usepackage{longtable}

\makeatletter
\AtBeginDocument{\@ifpackageloaded{rotating}{\PassOptionsToPackage{figuresright}{rotating}}{\usepackage[figuresright]{rotating}}\setlength{\rotFPtop}{0pt plus 1fil}\setlength{\rotFPbot}{0pt plus 1fil}}
\makeatother

\makeatletter
 \AtBeginDocument{%
  \@ifpackagewith{endfloat}{figuresonly}
  {\DeclareDelayedFloatFlavor{sidewaysfigure}{figure}}%true
  {\@ifpackagewith{endfloat}{tablesonly}{\DeclareDelayedFloatFlavor{sidewaystable}{table}\DeclareDelayedFloatFlavor{longtable}{table}\DeclareDelayedFloatFlavor{landscape}{table}}%true
  {\@ifpackageloaded{endfloat}{\DeclareDelayedFloatFlavor{sidewaysfigure}{figure}\DeclareDelayedFloatFlavor{sidewaystable}{table}\DeclareDelayedFloatFlavor{longtable}{table}\DeclareDelayedFloatFlavor{landscape}{table}}{}}%false
  }%false
  }
\makeatother

\usepackage{pdflscape}

\begin{document}

\begin{frontmatter}

\title{Review of Polymer Solutions for Near-Field Electrospinning with Spatial Control
}

\author[]{Antonio Osamu Katagiri Tanaka}
\ead{oskatagiri@gmail.com}
\author[]{H{\'e}ctor Al\'{a}n Aguirre Soto}
\ead{alan.aguirre@tec.mx}

\begin{abstract}
Near-field electrospinning (NFES) is identified to be a technique able to fabricate polymer nano and micro fibers with accurate placement. In the past years (2006-2019), several polymer solutions have been successfully electrospun into fibers through several variants of the conventional NFES process. Each NFES variant intents to tailor the process parameters in order to improve the fibers' properties. This paper presents a review on the research and related development of electrospun fibers, emphasizing the used polymers, solvents, and fiber characteristics. Relevant summary of polymer solutions and near-field electrospinning processing conditions is provided in this paper.
\end{abstract}
\begin{keyword} 
      polymer\sep solvent\sep near-field electrospinning\sep NFES\sep fibers\sep spatial control
\end{keyword}

\end{frontmatter}
\tableofcontents

\section{Introduction}
Even though electrospinning is an old invention \unskip~\cite{527120:12073288}, it is currently a trending topic among researchers \unskip~\cite{527120:12073453,527120:12073495,527120:12073496}. One of the reasons electrospinning is to be studied is its potential to fabricate polymer nano-fibers from a variety of polymers. The technique allows the production of thin continuous fibers with ease, with diameters down to 3 nm in some cases, which is something difficult to achieve by other techniques. Furthermore, the basic setup can be modified with ease to fabricate different fibers with diversified functionalities with different materials. The produced fibers can be aligned or unaligned. Besides, the electrospinning equipment is inexpensive and of small size, compared to the equipment of standard spinning techniques. On the other hand, the understanding of the electrospinning process has improved in the last years \unskip~\cite{527120:12073538}.

The main components of the electrospinning technique are the fluid control unit (e.g. syringe pump) and a voltage power supply. The process also requires a target electrode or combination of electrodes on which the fibers can be collected. Figure 6.2 describes a typical electrospinning set-up \unskip~\cite{527120:12073538}. Two sub-techniques can be derived from electrospinning depending on the distance between the dispensing electrode and the collector. The process in which the electrospun jet can be controlled near the tip is called NFES or near-field electrospinning \unskip~\cite{527120:12033655}. Moreover, if the distance between the collector and the dispensing needle is greater, the configuration is known as FFES or far-field electrospinning \unskip~\cite{527120:12073581}. Near-field electrospinning is considered to be an outstanding technique to fabricate polymer fibers with spatial control. NFES has suffered several modifications to improve the precision and accuracy of the fiber deposition.

\section{NFES}
\unskip~\cite{527120:11973130}

\unskip~\cite{527120:11974314}

\unskip~\cite{527120:11974329}

\unskip~\cite{527120:11974305}

\unskip~\cite{527120:11974306}

\unskip~\cite{527120:11974308}

\unskip~\cite{527120:11974310}

\unskip~\cite{527120:11974311}

\unskip~\cite{527120:11974312}

\unskip~\cite{527120:11974313}

\unskip~\cite{527120:11974316}

\unskip~\cite{527120:11974320}

\unskip~\cite{527120:11974321}

\unskip~\cite{527120:11974322}

\unskip~\cite{527120:11974323}

\unskip~\cite{527120:11974324}

\unskip~\cite{527120:11974325}

\unskip~\cite{527120:11974326}

\unskip~\cite{527120:11974327}

\unskip~\cite{527120:11974328}

\unskip~\cite{527120:11974304}

\unskip~\cite{527120:11974307}

\unskip~\cite{527120:12033655}

\unskip~\cite{527120:12033656}

\unskip~\cite{527120:12033657}

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

\begin{landscape}
\makeatletter\@twocolumnfalse\makeatother
\begingroup
\makeatletter\if@twocolumn\@ifundefined{theposttbl}{\gdef\TwoColDocument{true}\onecolumn\onecolumn}{}\fi\makeatother \setlength\LTcapwidth{\textheight}
\begin{longtable}{p{\dimexpr.1641\linewidth-2\tabcolsep}p{\dimexpr.13520000000000005\linewidth-2\tabcolsep}p{\dimexpr.1662\linewidth-2\tabcolsep}p{\dimexpr.46839999999999996\linewidth-2\tabcolsep}p{\dimexpr.0661\linewidth-2\tabcolsep}}
\caption{{Electrospun Polymer Solutions - Solution and Process Parameters} }
\label{tw-bab4042dace7}
\def\arraystretch{1}\\\endfirsthead \hline \noalign{\vskip3pt} \noalign{\textit{Table \thetable\ continued}} \noalign{\vskip3pt} \hline \endhead \hline \noalign{\vskip3pt} \noalign{\textit{\hfill Continued on next page}} \noalign{\vskip3pt} \endfoot \endlastfoot 
\tbltoprule Polymer(s) & Solvent(s) & NFES Variant & Process Parameters and Fiber Characterization & Ref.\\
\tblmidrule 
Poly(ethylene oxide) (PEO) &
  Deionized water &
  Low-Voltage NFES &
  \textbf{Solution Concentration:} 1, 2, and 3 $wt\% $ PEO \mbox{}\protect\newline \textbf{Nozzle:} 27 gauge type 304; stainless steel needle \mbox{}\protect\newline \textbf{Solution deposition rate:} lower than 1$\mu L / h $ \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} 1$mm $ \mbox{}\protect\newline \textbf{Substrate composition: }Pyrolyzed SU-8 carbon and Si \mbox{}\protect\newline \textbf{Applied voltage: }polymer jet initiated at 400-600 $V $ and dispensed at 200-400 $V $ \mbox{}\protect\newline \textbf{x-y stage velocity:} 10-40$mm/s $ \mbox{}\protect\newline \textbf{Fiber Diameter:} 50-425$nm $ \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} \textit{Not determined} &
  \unskip~\cite{527120:11973130}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly[2-methoxy-5-(2-ethylhexyloxy)-1,4-phenylenevinylene] (MEH-PPV) with Poly(ethylene oxide) (PEO) &
  acetonitrile toluene mixture (65/35); acetic acid toluene (17/83); pure toluene &
  Typical NFES process &
  \textbf{Solution Concentration:} \mbox{}\protect\newline 10$mg $ of MEH-PPV in 2$mL $ of toluene; 500$mL $ of MEH-PPV solution with 250$mg $ of PEO in 3.5$mL $ of acetonitrile; 500$mL $ of MEH-PPV solution with 250$mg $ of PEO in 3$mL $ of acetic acid / toluene (17 / 83). The resulting MEH-PPV/PEO concentration is 1:100 \mbox{}\protect\newline \textbf{Nozzle:} mm-diameter tip Tungsten spinneret in a 26 gauge needle \mbox{}\protect\newline \textbf{Solution deposition rate:} 50$\mu L / h $ \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} 500$\mu m $ \mbox{}\protect\newline \textbf{Substrate composition:} SiO2/Si (oxide thickness = 800 nm) \mbox{}\protect\newline \textbf{Applied voltage:} around 1.3$kV $ \mbox{}\protect\newline \textbf{x-y stage velocity:} 50$cm/s $ \mbox{}\protect\newline \textbf{Fiber Diameter:} 100$nm $ \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} around 100$\mu m $ &
  \unskip~\cite{527120:11974305}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(ethylene oxide) (PEO) &
  Water &
  Scanning Tip Electrospinning and NFES &
  \textbf{Solution Concentration:} 7$wt\% $ PEO \mbox{}\protect\newline \textbf{Nozzle:} Needle outer diameter of 200$\mu m $ and inner diameter of 100$\mu m $ \mbox{}\protect\newline \textbf{Solution deposition rate:} 0.1$\mu L / h $ \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} 500$\mu m $ \mbox{}\protect\newline \textbf{Substrate composition:} \textit{Not determined} \mbox{}\protect\newline \textbf{Applied voltage:} polymer jet initiated at 1.5 $kV $ and dispensed at 600$V $ \mbox{}\protect\newline \textbf{x-y stage velocity:} 120$mm/s $ \mbox{}\protect\newline \textbf{Fiber Diameter:} 709$\pm $131$nm $; 49-74$nm $ when applied voltage is 800$V $ \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} \textit{ Not determined} \mbox{}\protect\newline \textbf{Notes:} 108$m $ yield in 15$min $ with a fiber diameter of 709$\pm $131$nm $ &
  \unskip~\cite{527120:11974306}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(vinylidine fluorid) (PVDF) &
  N,N Dimethylformamide (DMF) &
  Helix Electrohydro-dynamic Printing (HE-printing) &
  \textbf{Solution Concentration:} 1.8$g $ PVDF in 4.1$g $ of DMF and 4.1$g $ of acetone. The resulting concentration is 18\% PVDF. \mbox{}\protect\newline \textbf{Nozzle:} Needle outer diameter of 510$\mu m $ and inner diameter of 260$\mu m $ \mbox{}\protect\newline \textbf{Solution deposition rate:} 400$nL/min $ \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} 10-50$mm $ \mbox{}\protect\newline \textbf{Substrate composition: }Poly(dimethylsiloxane) (PDMS) on Ecoflex \mbox{}\protect\newline \textbf{Applied voltage:} 1.5{\textendash}3$kV $ \mbox{}\protect\newline \textbf{x-y stage velocity:} 0-400$mm/min $ \mbox{}\protect\newline \textbf{Fiber Diameter:} about 1.5-3$\mu m $ \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} \textit{Not determined} &
  \unskip~\cite{527120:11974308}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Polyhedral Oligomeric Silsesquioxane-Poly(Carbonate-Urea)Urethane (POSS-PCU) and Polyhedral Oligomeric Silsesquioxane Poly(Caprolactone-Poly(Carbonate-Urea)Urethane) (POSS-PCL-PCU) &
  Dimethyl acetamide (DMAC) and 1-Butanol &
  Electrohydro-dynamic 3D Print-patterning or Electrohydro-dynamic Jetting &
  \textbf{Solution Concentration: }POSS-PCU and POSS-PCL-PCU used in 20\%$w/w $ concentration in DMAC \mbox{}\protect\newline \textbf{Nozzle:} needle of 750 $\mu m $ in diameter \mbox{}\protect\newline \textbf{Solution deposition rate:} less than 1$\mu L / min $ \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance: }about between 500$\mu m $ to 2$mm $ \mbox{}\protect\newline \textbf{Substrate composition:} \textit{Not determined} \mbox{}\protect\newline \textbf{Applied voltage:} 8.0-10.0$kV $ \mbox{}\protect\newline \textbf{x-y stage velocity:} 10$mm/s $ \mbox{}\protect\newline \textbf{Fiber Diameter:} 5-50$\mu m $ \mbox{}\protect\newline \textbf{Distance between adjacent fibers: }250$\mu m $ &
  \unskip~\cite{527120:11974310}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(ethylene oxide) (PEO) &
  Distilled water &
  Electrohydro-dynamic Writing or Mechanoelectrospinning (MES) &
  \textbf{Solution Concentration:} 6$wt\% $ PEO \mbox{}\protect\newline \textbf{Nozzle:} \textit{Not determined} \mbox{}\protect\newline \textbf{Solution deposition rate:} 1200$nL/min $ \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} 7.5$mm $ \mbox{}\protect\newline \textbf{Substrate composition:} \textit{Not determined} \mbox{}\protect\newline \textbf{Applied voltage:} polymer jet initiated at 2 $kV $ and dispensed at 0.8-1$kV $ \mbox{}\protect\newline \textbf{x-y stage velocity:} around 400$mm/s $ \mbox{}\protect\newline \textbf{Fiber Diameter:} 200-350$nm $ \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} 5$\mu m $ &
  \unskip~\cite{527120:11974311}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(ethylene oxide) (PEO) &
  Deionized water and the ethanol with a volume ratio of 3:1 &
  Airflow-assisted Electrohydro-dynamic Direct-writing (EDW) &
  \textbf{Solution Concentration:} 8$wt\% $ PEO \mbox{}\protect\newline \textbf{Nozzle:} Outer airflow passage diameter: 1$mm $ Airflow gas pump pressure: 25$kPa $ Inner liquid passage diameter: 0.21$mm $ \mbox{}\protect\newline \textbf{Solution deposition rate:} 30$\mu L / h $ \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} 2$mm $ \mbox{}\protect\newline \textbf{Substrate composition: }Silicon \mbox{}\protect\newline \textbf{Applied voltage:} about 2$kV $ \mbox{}\protect\newline \textbf{x-y stage velocity:} 1-20$mm/s $ \mbox{}\protect\newline \textbf{Fiber Diameter:} 3.73 $\pm $ 1.37$\mu m $ \mbox{}\protect\newline \textbf{Distance between adjacent fibers: }5.13 $\pm $ 6.67$\mu m $ &
  \unskip~\cite{527120:11974312}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(Vinylidene Fluoride) (PVDF) &
  Acetone and Dimethyl Sulfoxide (DMSO) &
  3D Electrospinning &
  \textbf{Solution Concentration:} 17$wt\% $ PVDF; 1.7$g $ of PVDF, 5$g $ of acetone, 0.5$g $ of Capstone FS-66, 5$g $ of DMSO \mbox{}\protect\newline \textbf{Nozzle:} Needle inner diameter of 100$\mu m $ \mbox{}\protect\newline \textbf{Solution deposition rate:} 14$\;nL/min $ \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} 750$\mu m $ \mbox{}\protect\newline \textbf{Substrate composition:} A4 size commercial printing paper (Double A) \mbox{}\protect\newline \textbf{Applied voltage:} 1.9$kV $ \mbox{}\protect\newline \textbf{x-y stage velocity:} 10$mm/s $ \mbox{}\protect\newline \textbf{Fiber Diameter:} \textit{Not determined} \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} \textit{Not determined} &
  \unskip~\cite{527120:11974313}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(9-Vinyl Carbazole) (PVK) &
  Styrene &
  Typical NFES process &
  \textbf{Solution Concentration:} 3.96$wt\% $ PVK in styrene \mbox{}\protect\newline \textbf{Nozzle:} Needle inner diameter of 100$\mu m $ \mbox{}\protect\newline \textbf{Solution deposition rate:} 500$nL/min $ \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} around 2.5$mm $ \mbox{}\protect\newline \textbf{Substrate composition:} Si/SiO2 \mbox{}\protect\newline \textbf{Applied voltage:} 3-4$kV $ \mbox{}\protect\newline x-y stage velocity: 13.3$cm/s $ \mbox{}\protect\newline \textbf{Fiber Diameter:} 289.26 $\pm $ 35.37$nm $ \mbox{}\protect\newline \textbf{Distance between adjacent fibers: }50$\mu m $ \mbox{}\protect\newline \textbf{Notes:} 15$m $ yield in 2$min $ &
  \unskip~\cite{527120:11974316}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Polystyrene (PS) &
  1,2,4-Trichloro benzene &
  Electrohydro-dynamic (EHD) jet printing &
  \textbf{Solution Concentration:} 1 to 5$wt\% $ PS \mbox{}\protect\newline \textbf{Nozzle:} Glass nozzle inner diameter of 2$\mu m $ and outer diameter of 2.66$\mu m $ \mbox{}\protect\newline \textbf{Solution deposition rate:} Si \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance: 20, 30, 40$\mu m $} \mbox{}\protect\newline \textbf{Substrate composition: } \mbox{}\protect\newline \textbf{Applied voltage:} 500 to 400$V $ in 25$V $ increments \mbox{}\protect\newline \textbf{x-y stage velocity:} 0.01-10$mm/s $ \mbox{}\protect\newline \textbf{Fiber Diameter:} about 60-170$\mu m $ \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} \textit{Not determined} &
  \unskip~\cite{527120:11974320}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(ethylene oxide) (PEO) &
  \textit{Not determined} &
  Typical NFES process &
  \textbf{Solution Concentration:} 3$wt\% $ PEO \mbox{}\protect\newline \textbf{Nozzle:} \textit{Not determined} \mbox{}\protect\newline \textbf{Solution deposition rate:} \textit{Not determined} \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} 500$\mu m $ \mbox{}\protect\newline \textbf{Substrate composition:} Si \mbox{}\protect\newline \textbf{Applied voltage:} 1000$V $ \mbox{}\protect\newline \textbf{x-y stage velocity:} 20$cm/s $ \mbox{}\protect\newline \textbf{Fiber Diameter:} 300$nm $ \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} 25$\mu m $ &
  \unskip~\cite{527120:11974321}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(ethylene oxide) (PEO) &
  Distilled water &
  Multinozzle NFES &
  \textbf{Solution Concentration:} 5$wt\% $ \mbox{}\protect\newline \textbf{Nozzle:} four-nozzle and six-nozzle array with needle spacing changes from 1.5$mm $ to 3.5$mm $ \mbox{}\protect\newline \textbf{Solution deposition rate:} 1-3$\mu L / min $ \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} 2$mm $ \mbox{}\protect\newline \textbf{Substrate composition:} \textit{Not determined} \mbox{}\protect\newline \textbf{Applied voltage:} 1.7-2.7$kV $ \mbox{}\protect\newline \textbf{x-y stage velocity:} \textit{Not determined} \mbox{}\protect\newline \textbf{Fiber Diameter:} 5.47$\mu m $ \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} 3-5 $mm $ &
  \unskip~\cite{527120:11974322}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(ethylene oxide) (PEO) &
  Distilled water &
  Multinozzle NFES &
  \textbf{Solution Concentration:}5$wt\% $ \mbox{}\protect\newline \textbf{Nozzle:} Dual-28G-needle array with needle inner diameter of 0.18$mm $ and outer diameter of 0.36$mm $; with needle spacing changes from 2.0$mm $ to 3.0$mm $ \mbox{}\protect\newline \textbf{Solution deposition rate:} 0.2$\mu L / min $ \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} 3.0-4.0$mm $ \mbox{}\protect\newline \textbf{Substrate composition: } \textit{Not determined} \mbox{}\protect\newline \textbf{Applied voltage:} 2.0-3.0$kV $ \mbox{}\protect\newline \textbf{x-y stage velocity:} 20$mm/s $ \mbox{}\protect\newline \textbf{Fiber Diameter:} \textit{Not determined} \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} 218-326$\mu m $ &
  \unskip~\cite{527120:11974323}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(ethylene oxide) (PEO) &
  Distilled water &
  Multinozzle NFES &
  \textbf{Solution Concentration:}5 $wt\% $ \mbox{}\protect\newline \textbf{Nozzle:} Dual-28G-needle array with needle inner diameter of 180$\mu m $ and outer diameter of 360$\mu m $; with needle spacing changes of 2.0$mm $ \mbox{}\protect\newline \textbf{Solution deposition rate:} 0.2$\mu L / min $ \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} 4.0$mm $ \mbox{}\protect\newline \textbf{Substrate composition:} chromium-plated glass \mbox{}\protect\newline \textbf{Applied voltage:} 2.5$kV $ \mbox{}\protect\newline \textbf{x-y stage velocity:} 20$mm/s $ \mbox{}\protect\newline \textbf{Fiber Diameter:} \textit{Not determined} \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} 2.3002-2.7224$mm $ &
  \unskip~\cite{527120:11974324}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(ethylene oxide) (PEO) &
  \textit{Not determined} &
  Typical NFES process &
  \textbf{Solution Concentration:} 2$wt\% $ \mbox{}\protect\newline \textbf{Nozzle:} G30 needle with inner diameter of 0.15$mm $ \mbox{}\protect\newline \textbf{Solution deposition rate:} \textit{Not determined} \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} 1-3$mm $ \mbox{}\protect\newline \textbf{Substrate composition:} Silicon \mbox{}\protect\newline \textbf{Applied voltage:} 1250$V $ \mbox{}\protect\newline \textbf{x-y stage velocity:} \textit{Not determined} \mbox{}\protect\newline \textbf{Fiber Diameter:} \textit{Not determined} \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} 20$\mu m $ &
  \unskip~\cite{527120:11974325}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Gelatin \mbox{}\protect\newline (porcine skin) &
  Acetic Acid and Ethyl Acetate &
  Typical NFES process &
  \textbf{Solution Concentration:} 11$wt\% $ gelatin, 30$wt\% $ water, 35.4$wt\% $ acetic acid, 23.6$wt\% $ ethyl acetate \mbox{}\protect\newline \textbf{Nozzle:} 19G needle tip with outer diameter of 1.08$mm $ \mbox{}\protect\newline \textbf{Solution deposition rate:} \textit{Not determined} \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} 1.25$mm $ \mbox{}\protect\newline \textbf{Substrate composition:} Poly(Dimethylsiloxane) (PDMS) films \mbox{}\protect\newline \textbf{Applied voltage:} 1000$V $ \mbox{}\protect\newline \textbf{x-y stage velocity:} \textit{Not determined} \mbox{}\protect\newline \textbf{Fiber Diameter:} around 2-3$\mu m $ \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} 40$\mu m $ &
  \unskip~\cite{527120:11974326}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(ethylene oxide) (PEO) &
  Water/Ethanol (v/v = 60/40) &
  Typical NFES process &
  \textbf{Solution Concentration:} PEO concentrations of 16\% adn 18\% \mbox{}\protect\newline \textbf{Nozzle:} 40$\mu m $ \mbox{}\protect\newline \textbf{Solution deposition rate:} \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} 1$mm $ \mbox{}\protect\newline \textbf{Substrate composition:} Planar silicon \mbox{}\protect\newline \textbf{Applied voltage:} 1.7$kV $ \mbox{}\protect\newline \textbf{x-y stage velocity:} 0.36$m/s $ \mbox{}\protect\newline \textbf{Fiber Diameter:} 5.15$\mu m $ \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} \textit{Not determined} &
  \unskip~\cite{527120:11974327}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(ethylene oxide) (PEO) &
  Water/Ethanol (v/v = 3/1) &
  Electrohydro-dynamic Direct-Write (EDW) &
  \textbf{Solution Concentration:} 14$wt\% $ PEO \mbox{}\protect\newline \textbf{Nozzle:} Stainless needle with inner diameter of 210$\mu m $ and outer diameter of 400$\mu m $ \mbox{}\protect\newline \textbf{Solution deposition rate:} 50$\mu L/h $ \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} 2$mm $ \mbox{}\protect\newline \textbf{Substrate composition:} Poly(ethylene terephthalate) (PET) \mbox{}\protect\newline \textbf{Applied voltage:} 3$kV $ \mbox{}\protect\newline \textbf{x-y stage velocity:} 700$mm/s $ \mbox{}\protect\newline \textbf{Fiber Diameter:} 15-35$\mu m $ \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} 70$\mu m $ &
  \unskip~\cite{527120:11974328}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(ethylene oxide) (PEO) &
  Deionized water &
  Mechano-Electrospinning &
  \textbf{Solution Concentration:} 3$wt\% $ PEO \mbox{}\protect\newline \textbf{Nozzle:} Stainless steel nozzle with inner diameter of 160$\mu m $ and outer diameter of 310$\mu m $ \mbox{}\protect\newline \textbf{Solution deposition rate:} 50$nL/min $ \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} 2-5$mm $ \mbox{}\protect\newline \textbf{Substrate composition:} Silicone \mbox{}\protect\newline \textbf{Applied voltage:}polymer jet initiated at 2$kV $ and dispensed at 1$kV $ \mbox{}\protect\newline \textbf{x-y stage velocity:} 200-400$mm/s $ \mbox{}\protect\newline \textbf{Fiber Diameter:} from 344$\pm $32 to 214$\pm $27$nm $ \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} \textit{Not determined} &
  \unskip~\cite{527120:11974304}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(co-Glycolic) acid (PLGA)  &
  Dimethyl Carbonate (DMC) &
  Tethered Pyro-Electrohydro-dynamic Spinning (TPES) &
  \textbf{Solution Concentration:} \textit{Not determined} \mbox{}\protect\newline \textbf{Nozzle:} nozzle-free \mbox{}\protect\newline \textbf{Solution deposition rate:} The drop reservoir is placed directly on a flat substrate \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} Taylor's cone is focused and put in direct contact with the collector \mbox{}\protect\newline \textbf{Substrate composition:} Poly(tetrafluoroethylene) (PTFE) coated glass slide \mbox{}\protect\newline \textbf{Applied voltage:} pyro-electric field of between 2.7 $x10^{7}\;V/m $ and 5.5$x10^{7}\;V/m $ \mbox{}\protect\newline \textbf{x-y stage velocity:} \textit{Not determined} \mbox{}\protect\newline \textbf{Fiber Diameter:} 304.7$nm $ \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} \textit{Not determined} &
  \unskip~\cite{527120:11974307}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(ethylene oxide) (PEO) with Tetrabutylammonium tetrafluoroborate (TBF) and SU-8 2002 &
  N,N Dimethylformamide (DMF) &
  Typical NFES process &
  \textbf{Solution Concentration:} SU-8/PEO/TBF blend with 0.75$wt\% $ PEO, 1$wt\% $ TBF; the blend is diluted with 30$vol\% $ DMF \mbox{}\protect\newline $\mu m $$\mu m $ \mbox{}\protect\newline \textbf{Solution deposition rate:} \textit{Not determined} \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} \textit{Not determined} \mbox{}\protect\newline \textbf{Substrate composition:} Brass disk with a diameter of 38$mm $ \mbox{}\protect\newline \textbf{Applied voltage:} 980$V $ \mbox{}\protect\newline \textbf{x-y stage velocity:} \textit{Not determined} \mbox{}\protect\newline \textbf{Fiber Diameter:} \textit{Not determined} \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} \textit{Not determined} &
  \unskip~\cite{527120:12033655}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(ethylene oxide) (PEO) &
  Water:Ethanol (3:2) &
  Suspension NFES &
  \textbf{Solution Concentration:} 14$wt\% $ PEO \mbox{}\protect\newline \textbf{Nozzle:} stainless steel needle (25 G) with inner diameter of 0.25$mm $ \mbox{}\protect\newline \textbf{Solution deposition rate:} 3$nL/s $ \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} between 0.5 and 10$mm $ with 0.5$mm $ increments \mbox{}\protect\newline \textbf{Substrate composition:} Planar silicon electrodes \mbox{}\protect\newline \textbf{Applied voltage:} 1.6$kV $ \mbox{}\protect\newline \textbf{x-y stage velocity:} 50, 150, and 250$mm/s $ \mbox{}\protect\newline \textbf{Fiber Diameter:} 300$nm $ \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} 0.1 and 0.5$mm $ &
  \unskip~\cite{527120:12033656}\\\cline{1-1}\cline{2-2}\cline{3-3}\cline{4-4}\cline{5-5}
Poly(ethylene oxide) (PEO) &
  Deionized water &
   &
  \textbf{Solution Concentration:} 10$wt\% $ PEO \mbox{}\protect\newline \textbf{Nozzle:} 32G metal needle \mbox{}\protect\newline \textbf{Solution deposition rate:} (Jet impact speed of 5$mm/s $ ) \mbox{}\protect\newline \textbf{Nozzle-to-substrate distance:} 0.5$mm $ \mbox{}\protect\newline \textbf{Substrate composition:} p-type silicon wafer \mbox{}\protect\newline \textbf{Applied voltage:} 400$V $ \mbox{}\protect\newline \textbf{x-y stage velocity:} 5$mm/s $ \mbox{}\protect\newline \textbf{Fiber Diameter:} \mbox{}\protect\newline \textbf{Distance between adjacent fibers:} 50$\mu m $ &
  \unskip~\cite{527120:12033657}\\
\tblbottomrule 
\end{longtable}
\endgroup
\makeatletter\@ifundefined{TwoColDocument}{}{\twocolumn}\makeatother 
\end{landscape}
Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\section{Polymer Solution and Process Parameters}

\section{Applications}

\section{Fiber Characterization}

\section{Conclusion}

\apptocmd{\thebibliography}{\csname phantomsection\endcsname\addcontentsline{toc}{section}{\numberline{}\refname}}{}{}

\bibliographystyle{elsarticle-num}

\bibliography{\jobname}

\end{document}