### 2020-05-28 myThesis/thesis_LaTeX/Chapters/4.tex

% Chapter Template

\chapter{Compatible Polymer-Solvent Combinations for Near-Field Electrospinning and Pyrolysis} % Main chapter title

\label{Chapter:4}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------
\section{Design and Selection of Candidate Spunable Polymer Solutions}

%high molecular weight, polymer-solution interaction (straight chains)

\subsection{Rheology of candidate polymer solutions}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------
\section{Effect of aromatic groups in oxygen-free polymers in NFES and Pyrolysis}