### 2020-05-28 myThesis/thesis_LaTeX/Chapters/6.tex

% Chapter Template

\chapter{Fabrication of Conductive Carbon nanofibers by Pyrolysis of Polymeric Fibers} % Main chapter title

\label{Chapter:6}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------
\section{}

%Yield rate
%Carbon content
%Conductivity
%Fiber morphology