### 2020-05-28 myThesis/thesis_LaTeX/Chapters/7.tex

% Chapter Template

\chapter{Comparison of the Carbon Fibers Obtained Against SU8-based Carbon Fibers} % Main chapter title

\label{Chapter:7}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------
\section{Fabrication and Characterization of Legacy SU-8 carbon fibers}