### 2020-06-04 myThesis/thesis_LaTeX/Chapters/4.tex

% Chapter Template

\chapter{Fabrication of Polymeric Fibers through Near-Field Electrospinning, Photopolymerization and Pyrolysis of Polymeric Fibers} % Main chapter title

\label{Chapter:4}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------
\section{}

%----------------------------------------------------------------------------------------
%	SECTION 2
%----------------------------------------------------------------------------------------
\section{}

%----------------------------------------------------------------------------------------
%	SECTION 3
%----------------------------------------------------------------------------------------
\section{Fabrication and Characterization of Legacy SU-8 carbon fibers}

\section{\emph{conclude with fibre morphology before and after pyrolysis. determine best pyrolysis process}}
%Were the selected candidates electrospunable?
%Fiber morphology
%Yield rate
%Carbon content
%Conductivity
%Fiber morphology