### 2020-06-14 myThesis/thesis_LaTeX/Chapters/5.tex
thesis characterization chapter added

% Chapter Template

\chapter{Comparison of the Obtained Carbon Fibres Against SU8-based Carbon Fibres and Potential Applications} % Main chapter title

\label{Chapter:5}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------
\section{Characterization \& Reproducibility}

%Were the selected candidates electrospunable?
%Fiber morphology