### 2020-08-20 5_Thesis/THESIS_LaTeX/Chapters/3.tex
file re organisation

% Chapter Template

\chapter{Selection of Compatible Polymer-Solvent Combinations for Near-Field Electrospinning and Pyrolysis} % Main chapter title

\label{Chapter:3}

% search : electrospunable capacity vs. thermo analysis of polymers in solutions (bad/good solvents)

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------
\section{Selection of Candidate Spunable Polymer Solutions}

%high molecular weight, polymer-solution interaction (straight chains)

\subsection{Rheology of candidate polymer solutions}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------
\section{Effect of aromatic groups in oxygen-free polymers in NFES and Pyrolysis}

\section{\emph{conclude with a collection of potential spunable polymer solutions}}