### 2020-11-07 05_Thesis/THESIS_LaTeX/Chapters/2.tex
able to fit lin

\chapter{Near-Field Electrospinning as an Affordable Way to Gain Spatial Control} % Main chapter title

\label{Chapter:2}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------
\section{Review of Polymer Solutions for NFES with Spatial Control}

\section{\emph{conclude with a NFES fabrication parameter baseline to yield the desired fibres}}