### 2020-11-07 05_Thesis/THESIS_LaTeX/Chapters/5.tex

% Chapter Template

\chapter{Concluding Remarks} % Main chapter title

\label{Chapter:5}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------
\section{}

\section{Future work}
% enhance conductivity with photopolymerization, CNT streching, ...