### 2020-11-20 05_Thesis/THESIS_LaTeX/Appendices/Appendix_ViscosityCurves.tex

% Appendix Template

\chapter{Flow Curves} % Main appendix title

\label{AppendixX} % Change X to a consecutive letter; for referencing this appendix elsewhere, use \ref{AppendixX}

\begin{figure}[!th]
\centering
\includegraphics[width=\textwidth]{./Figures/plt_PEOwtinSU8.png}
\decoRule
\caption[Viscosity as a function of shear rate for Poly(Ethylene Oxide) (PEO) and SU-8 2002 solutions]{Viscosity as a function of shear rate for Poly(Ethylene Oxide) (PEO) and SU-8 2002 solutions}
\label{fig:plt_PEOwtinSU8}
\end{figure}

\begin{figure}[!th]
\centering
\includegraphics[width=\textwidth]{./Figures/plt_PSwtinTHF.png}
\decoRule
\caption[Viscosity as a function of shear rate for Polystyrene (PS) and Tetrahydrofuran (THF) solutions]{Viscosity as a function of shear rate for Polystyrene (PS) and Tetrahydrofuran (THF) solutions}
\label{fig:plt_PSwtinTHF}
\end{figure}

\begin{figure}[!th]
\centering
\includegraphics[width=\textwidth]{./Figures/plt_PSBwtinNMP.png}
\decoRule
\caption[Viscosity as a function of shear rate for Poly(Styrene-co-Butadiene) (PSB) and 1-Methyl-2-Pyrrolidinone (NMP) solutions]{Viscosity as a function of shear rate for Poly(Styrene-co-Butadiene) (PSB) and 1-Methyl-2-Pyrrolidinone (NMP)}
\label{fig:plt_PSBwtinNMP}
\end{figure}

\begin{figure}[!th]
\centering
\includegraphics[width=\textwidth]{./Figures/plt_PSBwtinTHFandDMF.png}
\decoRule
\caption[Viscosity as a function of shear rate for Poly(Styrene-co-Butadiene) (PSB), Tetrahydrofuran (THF) and N,N-Dimethylformamide (DMF) solutions]{Viscosity as a function of shear rate for Poly(Styrene-co-Butadiene) (PSB), Tetrahydrofuran (THF) and N,N-Dimethylformamide (DMF) solutions}
\label{fig:plt_PSBwtinTHFandDMF}
\end{figure}

\begin{figure}[!th]
\centering
\includegraphics[width=\textwidth]{./Figures/plt_PSMSwtinDMF.png}
\decoRule
\caption[Viscosity as a function of shear rate for Poly(Styrene-co-alpha-Methylstyrene) (PSMS) and N,N-Dimethylformamide (DMF) solutions]{Viscosity as a function of shear rate for Poly(Styrene-co-alpha-Methylstyrene) (PSMS) and N,N-Dimethylformamide (DMF) solutions}
\label{fig:plt_PSMSwtinDMF}
\end{figure}

\begin{figure}[!th]
\centering
\includegraphics[width=\textwidth]{./Figures/plt_PVKwtinCHL.png}
\decoRule
\caption[Viscosity as a function of shear rate for Poly(9-Vinylcarbazole) (PVK) and Chloroform (CHL) solutions]{Viscosity as a function of shear rate for Poly(9-Vinylcarbazole) (PVK) and Chloroform (CHL) solutions}
\label{fig:plt_PVKwtinCHL}
\end{figure}

\begin{figure}[!th]
\centering
\includegraphics[width=\textwidth]{./Figures/plt_PVKwtinSU8.png}
\decoRule
\caption[Viscosity as a function of shear rate for Poly(9-Vinylcarbazole) (PVK) and SU-8 2002 solutions]{Viscosity as a function of shear rate for Poly(9-Vinylcarbazole) (PVK) and SU-8 2002 solutions}
\label{fig:plt_PVKwtinSU8}
\end{figure}