### 2020-11-21 05_Thesis/THESIS_LaTeX/Appendices/Appendix_MicroscopyChar.tex
file updates

% Appendix Template

\chapter{Optical Microscopy Characterization of Electrospun Fibers} % Main appendix title

\label{Appendix_MicroscopyChar} % Change X to a consecutive letter; for referencing this appendix elsewhere, use \ref{AppendixX}

\begin{figure}[!th]
\centering
\includegraphics[scale=0.55]{./Figures/microscopy_peo_su8.png}
\decoRule
\caption[Morphology and Characterization of Electrospun Fibers at Different Voltages : Poly(Ethylene Oxide) (PEO) and SU-8 2002 ]{Morphology and Characterization of Electrospun Fibers at Different Voltages : Poly(Ethylene Oxide) (PEO) and SU-8 2002 }
\label{fig:microscopy_peo_su8}
\end{figure}

\begin{figure}[!th]
\centering
\includegraphics[scale=0.55]{./Figures/microscopy_ps_thf.png}
\decoRule
\caption[Morphology and Characterization of Electrospun Fibers at Different Voltages : Polystyrene (PS) in Tetrahydrofuran (THF) ]{Morphology and Characterization of Electrospun Fibers at Different Voltages : Polystyrene (PS) in Tetrahydrofuran (THF) }
\label{fig:microscopy_ps_thf}
\end{figure}

\begin{figure}[!th]
\centering
\includegraphics[scale=0.55]{./Figures/microscopy_psb_thf_dmf.png}
\decoRule
\caption[Morphology and Characterization of Electrospun Fibers at Different Voltages : Poly(Styrene-co-Butadiene) (PSB) in Tetrahydrofuran (THF) and N,N-Dimethylformamide (DMF) ]{Morphology and Characterization of Electrospun Fibers at Different Voltages : Poly(Styrene-co-Butadiene) (PSB) in Tetrahydrofuran (THF) and N,N-Dimethylformamide (DMF) }
\label{fig:microscopy_psb_thf_dmf}
\end{figure}

\begin{figure}[!th]
\centering
\includegraphics[scale=0.55]{./Figures/microscopy_pvk_chl.png}
\decoRule
\caption[Morphology and Characterization of Electrospun Fibers at Different Voltages : Poly(9-Vinylcarbazole) (PVK) in Chloroform (CHL) ]{Morphology and Characterization of Electrospun Fibers at Different Voltages : Poly(9-Vinylcarbazole) (PVK) in Chloroform (CHL) }
\label{fig:microscopy_pvk_chl}
\end{figure}

\begin{figure}[!th]
\centering
\includegraphics[scale=0.55]{./Figures/microscopy_pvk_su8.png}
\decoRule
\caption[Morphology and Characterization of Electrospun Fibers at Different Voltages : Poly(9-Vinylcarbazole) (PVK) and SU-8 2002 ]{Morphology and Characterization of Electrospun Fibers at Different Voltages : Poly(9-Vinylcarbazole) (PVK) and SU-8 2002 }
\label{fig:microscopy_pvk_su8}
\end{figure}