### 2020-12-02 05_Thesis/THESIS_LaTeX/Appendices/Appendix_ReviewTable.tex

% Appendix Template
\chapter{Appendix NFES Review Table} % Main appendix title

\label{Appendix_NFESreviewTable} % Change X to a consecutive letter; for referencing this appendix elsewhere, use \ref{AppendixX}

\begin{landscape}
\makeatletter\@twocolumnfalse\makeatother
\begingroup
\makeatletter\if@twocolumn\@ifundefined{theposttbl}{\gdef\TwoColDocument{true}\onecolumn\onecolumn}{}\fi\makeatother \setlength\LTcapwidth{\textheight}
{\tiny
\begin{longtable}{p{.075\textwidth} p{.025\textwidth} p{.080\textwidth} p{.05\textwidth} p{.05\textwidth} p{.05\textwidth} p{.05\textwidth} p{.05\textwidth} p{.05\textwidth} p{.05\textwidth} p{.05\textwidth} p{.05\textwidth} p{.05\textwidth} p{.05\textwidth} p{.05\textwidth} p{.05\textwidth} p{.05\textwidth} p{.05\textwidth} p{.05\textwidth} p{.05\textwidth} }
\caption{{Electrospun Polymer Solutions - Solution and Process Parameters} }
\def\arraystretch{1}\\\endfirsthead \hline \noalign{\vskip3pt} \noalign{\textit{Table \thetable\ continued}} \noalign{\vskip3pt} \hline \endhead \hline \noalign{\vskip3pt} \noalign{\textit{\hfill Continued on next page}} \noalign{\vskip3pt} \endfoot \endlastfoot 
Polymer & $Mw$ & Solvent & $\eta_0$ [Pa s] & $K$ [S / m] & $\gamma$ [N / m] & $\rho$ [Kg / $m^3$] & NFES Type & $C_{polymer}$ [wt$\%$] & $D_{nozzle}$ [m] & $Q$ [$m^3$ / s] & Substrate & $L$ [m] & $\Phi_0$ [V] & $v_{stage}$ [m / s] & $D_{fiber}$ [m] & $FiberGap$ [m] & Fiber Morphology & Fiber Structure & Reference \\
\hline
PEO / SU8 & 4e6 & Cyclopentanote & 0.2 - 32e-3 & 8e-3 - 11e-3 & 25 & 1149 - 1318 & xy stage & 0.25 - 1 & 108e-6 &  & Si & 0.1e-3 & 200 &  & 1.2e-6 - 3.49e-6 &  & straight & single & \cite{Aguirre-soto2002} \\
PCL & 80e3 & DCM / Methanol & 7e-3 &  &  & 1218.538 & xy stage & 8  -  20 & 337e-6 & 2.83e-10 & Al &  & 10000 - 20000 &  & 374e-9 - 1.33e-6 &  & straight & cross aligned & \cite{Beachley2011} \\
PCL & 50e6 & CF / DMF & 0.747 &  &  & 1142.62 & melt &  & 514e-6 & 1.39e-11 &  & 0.03 & 12000 & 4.5e-3 - 0.1e-3 & 10.6e-6 - 46.5e-6 &  & coil and straight & single & \cite{Brown2011} \\
PCL & 80e3 & CF / DMF & 747e-3 &  &  & 1142.62 & melt &  & 743e-6 & 1.39e-12 - 5.56e-12 & Al & 0.02 - 0.04 & 6000 - 12000 &  & 4.76e-6 - 33e-6 &  & coil & randomly distributed & \cite{Brown2014} \\
PEO & 300e3 & Water & 0.95 & 400e-6 & 1825.5 & 1000.6 & xy stage & 3 & 0.16e-3 & 8.33e-13 & Si & 0.005 & 1000 & 0.05 - 0.5 & 214e-9 - 344e-9 & 4.21e-5 & meander and straight & single & \cite{Bu2012} \\
MEH - PPV / PEO & 380e3 & Toluene & 573e-6 & 956.e-3 & 17e-3 & 865 & xy stage & 0.08 & 0.26e-3 & 1.39e-11 & SiO2 / Si & 0.5e-3 & 1300 & 0.5 & 1.38e-7 & 8.84e-5 & straight & aligned & \cite{Camillo2013} \\
PEO / SU8 & 4e6 & Cyclopentanote & 0.056 & 9e-3 & 25 & 1215.5 & xy stage & 0.25 & 108e-6 &  & Si & 0.1e-3 & 200 & 0.06 & 4.08e-6 &  & straight & single & \cite{Cardenas2017} \\
PEO & 300e6 & Water & 0.95 & 5.7e-3 & 8886.7 & 1000.6 & xy stage & 3 - 9 & 0.0001 & 2.80E - 14 &  & 0.0005  -  0.001 & 500  -  1200 & 0.12 & 119e-9 - 2.31e-6 &  & straight & mesh & \cite{Chang2008} \\
PVDF & 534e6 & CF / DMF & 0.74 &  &  & 1142.62 & xy stage & 18 &  &  &  & 0.35e-3 &  &  & 1.07e-6 - 2.16e-6 &  & straight & single & \cite{Chang2010} \\
PVDF & 534e3 & CF / DMF & 0.74 &  &  & 1142.62 & xy stage & 18 &  &  &  & 0.35e-3 &  &  & 6.13e-6 - 9.19e-6 &  & straight & aligned & \cite{Chang2011} \\
PU &  & CF / DMF & 0.74 &  &  & 1142.62 & xy stage & 10 &  &  & acrylic &  & 2000 & 0.1 & 9.32E - 06 & 5.22e-5 & straight & aligned & \cite{Choi2017} \\
PLGA &  & CF / DMF & 0.74 &  &  & 1142.62 & PYES &  &  &  & PTFE &  &  &  & 3.5e-7 &  & straight & aligned & \cite{Coppola2014} \\
PLGA & 54e3 & CF / DMF & 0.74 &  &  & 1142.62 & PYES & 25 &  &  & Si & 0.001 &  &  & 0.01e-3 - 0.03e-3 &  & straight & aligned & \cite{Coppola2020} \\
PCL & 83e3 &  &  &  &  &  & melt &  & 0.18e-3 &  & Al & 1.5e-3 & 2900 & 91.67e-3 & 8.17e-7 & 100.6e-6 & straight & mesh & \cite{Dalton2015} \\
PAN &  & CF / DMF & 0.74 &  &  & 1142.62 & rotating drum & 9 &  &  & Si & 1.85e-3 & 35 - 500 &  & 0.13e-6  - 0.349e-6 &  & straight & aligned & \cite{Deng2020} \\
PDO &  & CF / DMF & 0.0162  -  0.74 &  &  & 1142.62 - 1484 & xy stage & 14  -  20 & 340e-6  - 840e-6 & 2.78e-12 - 25.4e-12 & Al & 1.8e-3 - 0.003 & 1200 - 1600 & 0.01 - 0.1 & 3.19e-6 - 25.2e-6 &  & straight & mesh & \cite{E.King2019} \\
PMMA &  & CF / DMF & 0.74 &  &  & 1142.62 & xy stage & 16  -  24 &  & 13.9e-12 - 41.7e-12 &  & 0.0016 - 2.4e-3 & 1200 - 2400 & 0.06 & 1.41e-6 - 5.64e-6 &  & straight & mesh & \cite{Fattahi2017} \\
PEO / SU8 & 4e-6 & CF / DMF & 0.74 &  &  & 1142.62 & xy stage & 0.5 &  &  &  &  &  & 0.02  -  0.06 & 5.98E - 06 &  & straight & aligned & \cite{Flores2017} \\
PEO & 300e3 & Water & 0.95 & 859e-6 & 79e3 & 1e3 & xy stage & 8 & 0.4e-3 &  & Si & 0.5e-3 & 1100 & 0.2 & 7.4e-7 & 2e-5 & straight & single & \cite{Fuh2011} \\
PEO / Chitosan & 900e3 & DMSO & 2.1e-3 & 62e-3 & 4.48 & 1095.6 & xy stage & 5 & 0.4e-3 & 2.8e-14 &  & 0.5e-3 - 1e-3 & 800 - 1200 &  & 282e-12 - 785e-12 & 0.1e-3 & curved & aligned & \cite{Fuh2012} \\
PEO / Chitosan & 900e3 & CF / DMF & 0.74 &  &  & 1142.62 & xy stage & 5 &  &  & PPy & 0.5e-3 & 800 & 0.2 & 7.3e-7 & 0.02e-3 - 0.1e-3 & straight & mesh, randomly distributed and aligned & \cite{Fuh2013} \\
PVDF & 534e-3 & CF / DMF & 0.74 &  &  & 1142.62 &  & 16 &  &  &  & 1.2e-3 & 1440 &  & 907e-9 - 2.91e-6 &  & straight & aligned & \cite{Fuh2015} \\
PEO & 300e-3 & Water / Ethanol & 1.048316 & 2.6e-3 - 7.8e-3 & 11886.69 & 894.31 & xy stage & 20 & 4e-5 &  & Si & 1e-3 & 1700 & 0.05 - 0.35 & 816e-9 - 2.82e-6 &  & meander, coil and straight & single & \cite{GaofengZheng2010} \\
PAN & 150e3 & DMF & 0.94 &  &  & 802 & rotating drum & 11 & 159e-6 &  & Si & 0.7e-3 - 0.01 & 400 - 900 &  & 2.53e-6 &  & straight & aligned & \cite{George2020} \\
POSS - PCL - PCU & 2000 & CF / DMF & 0.74 & 10.2e-3 & 32235.64 & 1142.62 & xy stage & 20 & 750e-6 & 1.67e-11 &  & 0.5e-3 & 9000 & 0.01 & 18.7e-6 - 690e-6 & 352e-6 - 0.48e-3 & meander & mesh & \cite{Gupta2007} \\
PCL & 80e3 & CF / DMF & 0.74 &  &  & 1142.62 & xy stage &  & 912e-6 &  &  & 0.005 & 3000 & 0.05 & 108.3e6 - 150.5e-6 & 0.001 & straight & mesh & \cite{He2018} \\
PEtOx & 50e3 &  &  &  &  &  & melt &  & 160e-6 - 337e-6 &  & Al & 0.003  -  0.007 & 3000  -  7000 & 3.33e-3 & 12.5e-6 - 59.1e-6 &  & straight & cross aligned & \cite{Hochleitner2014} \\
PLA / PEG & 72000 &  &  &  &  &  & melt &  & 337e-6 &  & Si & 3.5e-3 & 4000 & 83.3e-3 & 3.1e-5 &  & straight & mesh & \cite{Hochleitner2017} \\
PEO & 300e3 & CF / DMF & 0.744 & 6.12e-5 & 34838.96 & 1142.62 & xy stage & 6 &  & 2e-11 &  & 0.005 - 7.5e-3 & 900 & 0.05 - 0.4 & 212e-9 - 352e-9 & 5e-6 & straight, coil, and meander & single and aligned & \cite{Huang2015} \\
PEO & 30000 & Water / Ethanol & 1.04 & 524.3e-6 & 11906.09 & 894.31 & xy stage & 8 & 210e-6 & 8.33e-12 & Si & 2e-3 & 2000 & 0.001 - 0.02 & 3.27e-6 - 9.77e-6 &  & meander & aligned & \cite{Jiang2018} \\
PEO & 100e3 & CF / DMF & 0.74 & 736.4e-6 & 4729.5 & 1142.62 & rotating drum & 10 &  &  & Si & 7.56e-3 - 12.5e-3 & 2000 & 0.11 - 1.70 & 101e-9 - 201e-9 &  & straight & aligned & \cite{Kameoka2003} \\
PEO & 900e3 & CF / DMF & 0.77 &  &  & 1142.62 & xy stage & 5 - 30 &  &  &  &  & 4500 &  & 150e-9 &  & coil & randomly distributed & \cite{Kameoka2003a} \\
PVDF & 534e3 & Acetone / DMSO & 0.16 &  &  & 943.42 & auxiliary electrodes & 17 & 0.1e-3 & 3.89e-12 & Paper & 750e-6 & 1900 & 0.01 &  &  & meander & single & \cite{Kim2018} \\
PCL & 50e3 & CF & 0.55 &  &  & 1483.24 & xy stage & 18.3 & 250e-6 &  &  &  & 6000 - 12000 & 5e-3  -  7.5e-3 & 33.4e-6 - 103e-6 &  & coil & single & \cite{Kolan2018} \\
PCL & 80e3 & CF & 0.55 & 997e-6 & 1978.8 & 1483.24 & xy stage & 8.8 & 0.2e-3 & 2.78e-11 & Si & 0.07 & 25e3 & 2e-3 - 0.5 &  &  & coil and meander & single & \cite{Lee2012} \\
PVDF & 534e3 & CF / DMF & 0.74 & 1.47e-5 & 4733.12 & 1142.62 & xy stage & 12 &  & 2.22E - 11 & Si & 5e-3 & 1200 & 0.03 & 4.23e-5 & 263.1e-6 & curved & mesh & \cite{Lei2015} \\
PVP & 1300e3 & CF / DMF & 0.73 & 2.94e-5 & 4737.7 & 1142.62 & xy stage & 17.4 & 510e-6 & 8.33e-11 & Si / SiO2 & 1e-3 & 50 & 0.8 & 1.7e-7 &  & straight & mesh & \cite{Lim2019} \\
PVDF & 534e-3 & Acetone / DMSO & 0.162 &  &  & 943.42 & xy stage & 18 & 0.3e-3 &  &  &  & 900  -  1200 & 0.03  -  0.09 & 2.54e-6 - 89.6e-6  &  & straight & aligned & \cite{Liu2013} \\
PVDF & 534e-3 & Acetone / DMSO & 0.162 &  &  & 943.42 & rotating drum & 18 & 381e-6 &  & Si & 0.5e-3 & 10e3 - 16e3 &  & 205e-9 - 1.45e-6 &  & straight & cross aligned & \cite{Liu2014} \\
PVDF & 534e3 & Acetone / DMSO & 0.16 &  &  & 943.42 & xy stage & 16  -  20 & 260e-6 & 1.67e-11 & Si & 1e-3 &  & 0.07 & 0.5e-6 - 20e-6 &  & straight & cross aligned & \cite{Liu2015a} \\
PVDF / PVA / SWCNT & 534e3 & Acetone / DMSO & 0.1 & 0.003 & 270 & 94 & xy stage & 21 & 159e-6 &  & paper & 0.001 & 1000  -  1800 & 0.01  -  0.09 & 3.99e-6  -  11.5e-6 & 0.1e-3 & straight & aligned & \cite{Luo2016} \\
PEO & 4e6 & Water & 0.95 & 0.57e-3 & 8892.6 & 1000.62 & xy stage & 2 & 210e-6 & 2.78e-13 & C /Si & 0.001 & 200  -  400 & 0.02 - 0.08 & 50.7e-9 - 423e-9 &  & straight & aligned and single & \cite{Madou2011} \\
PVK & 1100e3 & Styrene & 696e-6 & 6.31 & 3.6e-3 & 1300 & xy stage & 3.96 & 0.1e-3 & 8.33e-12 & Si / SiO2 & 2.5e-3 & 3500 & 133e-3 - 0.15 & 65e-9 - 707e-9 & 5e-5 & straight & aligned & \cite{Min2013} \\
PEO & 200e3 & Water / Ethanol & 1.04 & 523e-6 & 11907.5 & 894.3 & xy stage & 14 & 250e-6 & 3e-12 & Si & 5.25e-3 & 1600 & 0.15 & 3e-7 & 0.3e-3 & meander & aligned & \cite{Nagle2019} \\
PVDF & 534e3 & CF / DMF & 747e-3 &  &  & 1142.62 & rotating drum & 10  -  18 &  &  & Cu &  & 2000  -  10000 & 523.6e-3  -  2.6 & 35.3e-6  -  51.9e-6 &  & straight & aligned & \cite{Pan2014} \\
PVDF & 534e3 & CF / DMF & 747e-3 &  &  & 1142.62 & coaxial & 15  -  19 & 630e-6 &  & Cu &  & 8000  -  16000 & 0.53 - 2.62 & 10.3e-6 - 45.9e-6 &  & straight & aligned & \cite{Pan2015} \\
PCL & 45028 & CF / DMF & 0.74 &  &  & 1142.62 & xy stage & 25  -  70 & 0.3e-3 - 0.6e-3 & 0  -  1.21e-9 &  & 5e-3  -  0.01 & 3500  -  3900 & 0.2 & 58.7e-6 - 190e-6 &  & straight & single & \cite{Parajuli2016} \\
PEO / NaCl &  & CF / DMF & 0.7474335 &  &  & 1142.62 & xy stage & 0.1  -  1 &  &  &  &  &  &  & 92e-9  -  239e-9 &  & straight & aligned & \cite{Park2020} \\
PEO & 1000e3 & Water / Ethanol & 1.04 & 339e-6 & 11923.7 & 894.1 & xy stage & 4 & 160e-6 & 83.3e-12 - 25e-9 &  & 0.003 & 1500 &  & 5.74e-6  -  14.7e-6 &  & straight & mesh & \cite{Ru2014} \\
PEO / TiO2 & 300e3 & CF / DMF & 747e-3 & 618e-6 & 363503.0357 & 1142.62 & xy stage & 5.44 &  & 5.56e-12 & Si3N4 & 0.5e-3 & 1100 &  & 5e-7 & 4e-5 & straight & aligned & \cite{Ruggieri2013} \\
PIB &  &  &  &  &  &  & rotating drum & 15 & 360e-6 & 2e-11 &  &  & 12000 &  & 4.2e-7 &  & straight & cross aligned & \cite{Sarkar2007} \\
PS &  &  &  &  &  &  & rotating drum & 20 & 0.36e-3 & 1e-11 &  &  & 11000 &  & 4.1-7 &  & straight & cross aligned & \cite{Sarkar2007} \\
PEO / FeAl NP &  &  &  &  &  &  & rotating drum & 6 & 0.36e-3 & 1.17e-11 &  &  & 9500 &  & 4e-7 &  & straight & cross aligned & \cite{Sarkar2007} \\
PEO / Au NP &  &  &  &  &  &  & rotating drum & 10 & 0.36e-3 & 1.67e-11 &  &  & 9300 &  & 3.9e-7 &  & straight & cross aligned & \cite{Sarkar2007} \\
PEO / SWCNT &  &  &  &  &  &  & rotating drum & 8 & 360e-6 & 1.33e-11 &  &  & 9300 &  & 3.8e-7 &  & straight & cross aligned & \cite{Sarkar2007} \\
PEO &  &  &  &  &  &  & rotating drum & 8 & 360e-6 & 1.33e-11 &  &  & 9500 &  & 3.7e-7 &  & straight & cross aligned & \cite{Sarkar2007} \\
PEO / PPy & 5e6 & DMF / Acetone & 633.4e-3 &  &  & 948.8 & xy stage & 8 & 2.5e-5 &  &  &  & 1000 &  & 0.4e-6  -  0.7e-6 & 265e-6 & straight & aligned & \cite{Shen2016} \\
PEO & 400e3 & Water & 958.2e-3 & 286e-6 - 11e-3 & 2225.7 - 35611.8588 & 1000.62 & xy stage & 10 & 108e-6 &  &  & 750e-6 & 500  -  700 & 15e-3 & 130e-9  -  220e-9 &  & straight, coil and meander & single & \cite{Shin2018} \\
PEO & 4000e3 & Water & 958e-3 &  &  & 1000.62 & xy stage & 1.5 & 108e-6 &  &  &  & 650 & 0.2e-3 - 0.8e-3 & 94e-9  -  933e-9 &  & meander, coil and meander & single & \cite{Shin2019} \\
PS &  & TCB & 2.05e-3 &  &  & 1449.6 & xy stage & 5 & 2e-6 &  & Si & 20e-6  -  40e-6 & 400  -  500 & 2e-3 & 57.3e-9  -  167e-9 &  & straight & single & \cite{Song2015} \\
PCL & 80e3 & CF / DMF & 0.73 & 735e-6 & 4729.9 & 1142.62 & aligned needles & 10 & 337e-6 & 1.39e-10 & Si & 0.06 & 15000 &  & 1.21e-7 &  & straight & randomly distributed & \cite{Sonntag2020} \\
PEU &  & CF / DMF & 747e-3 &  &  & 1142.62 & touch spinning & 5 &  & 8.33e-11 &  &  &  &  & 3.28e-7 &  & straight & cross aligned & \cite{Strauss2019} \\
PCL &  & CF / DMF & 747e-3 &  &  & 1142.62 & touch spinning & 8 &  & 8.33e-11 &  &  &  &  & 3.26e-7 &  & straight & cross aligned & \cite{Strauss2019} \\
PEO / NT &  & CF / DMF & 0.74 &  &  & 1142.62 & touch spinning & 0.5 &  & 8.33e-11 &  &  &  &  & 3.2e-7 &  & straight & cross aligned & \cite{Strauss2019} \\
PEO &  & CF / DMF & 0.7474335 &  &  & 1142.62 & touch spinning & 0.5 &  & 8.33e-11 &  &  &  &  & 2.98e-7 &  & straight & cross aligned & \cite{Strauss2019} \\
PEO & 300e3 & CF / DMF & 0.7 &  &  & 1142.2 & xy stage & 3 &  &  & Si and SiO2 & 0.5e-3  -  3e-3 & 598  -  1255 & 0.2 & 3e-7 & 2.5e-5 & meander & single & \cite{Sun2006a} \\
PEO & 2000e3 & CF / DMF & 0.72 &  &  & 1142.6 & multinozzle xy stage & 5 &  & 3.33-11 &  & 2e-3  -  6e-3 & 1700  -  2700 &  &  & 5.47e-3 & straight & aligned & \cite{Wang2015} \\
PEO & 2e6 & Water & 0.92 &  &  & 1000.62 & multinozzle xy stage & 5 & 180e-6 & 3.33e-12 &  & 3e-3  -  5e-3 & 2300  -  2800 & 0.02 &  & 3.03e-3 & straight & aligned & \cite{Wang2017} \\
PEO & 2e6 & Water & 0.95 & 574.931e-3 & 8906.09 & 1000.62 & multinozzle xy stage & 5 & 180e-6 &  & SST & 4e-3 & 2500 & 0.02 & 6.23e-6 & 0.025 & straight & aligned & \cite{Wang2018a} \\
PEO & 4e6 &  &  &  &  &  & xy stage & 2 & 150e-6 &  & Si & 2e-3 & 1250 &  &  & 7e-6  -  74e-6 & meander & aligned & \cite{Xu2014} \\
PEO & 2e-6 & CF / DMF & 0.74 &  &  & 1142.62 & xy stage & 5 &  & 3.33e-12 &  & 2.5e-3  -  4e-3 & 1900  -  2700 & 0.05  -  0.3 & 696e-9  -  5.08e-6 &  & meander & aligned & \cite{Xu2019} \\
Gelatin &  & CF / DMF & 0.74 &  &  & 1142.62 & xy stage & 11 &  &  & PDMS & 1.25e-3 & 1000 &  & 9.02e-7 & 4e-5 & straight & aligned & \cite{Xue2014} \\
PEO & 900e3 & Water & 0.98 & 572.82e-6 & 8923.275983 & 1000.62 & xy stage & 5 & 110e-6 & 833e-15  -  2.78e-12 & PS & 0.5e-3  -  1.5e-3 & 1000  -  1300 & 8.77e-3  -  0.166 & 671e-9  -  19.9e-6 &  & straight & aligned & \cite{Yang2019} \\
PCL & 80e3 &  &  &  &  &  & xy stage &  & 260e-6 & 833e-15  -  2.78e-12 & PE & 2e-3 & 3000 &  & 6.8e-6 &  & meander & aligned & \cite{You2017} \\
PS & 240e3 & CF / DMF & 0.74 & 3.08e-3 & 910845.5 & 1142.62 & xy stage & 30 & 455e-6 & 0 & Au & 0.2e-3 & 2400 & 15e-3 & 7.81e-6 &  & straight & mesh & \cite{Yousefi2019} \\
PEO & 2e6 & CF / DMF & 0.75 & 735e-6 & 4749.86 & 1142.62 & multinozzle xy stage & 8 &  &  &  & 3e-3 & 2500 & 80e-3 & 2.89e-6  -  10e-6 & 2.06e-3 & straight & aligned & \cite{Zhang2019} \\
PEO & 300e3 & Water / Ethanol & 1.04 & 10.48e-3 & 11917.213 & 894.313 & xy stage & 18 & 4e-5 &  & Si & 1e-3 & 1700 & 0.36 & 4.89e-7 &  & coil & single & \cite{Zheng2010} \\
PEO & 300e3 & Water / Ethanol & 1.04 & 524e-6 & 11910.45 & 894.3 & xy stage & 14 & 210e-6 & 1.39e-11 & PET & 2e-3 & 500  -  4000 & 0.02  -  1 & 17.4e-6  -  65.8e-6 & 0.3e-3 & meander & aligned & \cite{Zheng2014} \\
PCL & 90e3 & CF / DMF & 0.74 & 7.37e-5 & 4735.564517 & 1142.62 & xy stage & 10 & 410e-6 & 16.7e-12  -  83.3e-12 & Si / Al & 1e-3  -  3e-3 & 800  -  4000 & 0.03 & 7.99e-6  -  91.5e-6 &  & straight & single & \cite{Zhou2011} \\
PEO & 2e6 & Water & 0.958 & 572e-6 & 8896.48 & 1000.62 & auxiliary electrodes & 4 & 160e-6 &  & Si & 4e-3 & 0  -  5000 & 5e-3 & 7.06e-6  -  13.6e-6 &  & meander and straight & single & \cite{Zhu2016} \\
\end{longtable}
}
\endgroup
\makeatletter\@ifundefined{TwoColDocument}{}{\twocolumn}\makeatother 
\end{landscape}