### 2020-12-02 05_Thesis/THESIS_LaTeX/main.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Masters/Doctoral Thesis 
% LaTeX Template
% Version 2.5 (27/8/17)
%
% This template was downloaded from:
% http://www.LaTeXTemplates.com
%
% Version 2.x major modifications by:
% Vel (vel@latextemplates.com)
%
% This template is based on a template by:
% Steve Gunn (http://users.ecs.soton.ac.uk/srg/softwaretools/document/templates/)
% Sunil Patel (http://www.sunilpatel.co.uk/thesis-template/)
%
% Template license:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
% The default document font size, options: 10pt, 11pt, 12pt
11pt, 
% Two side (alternating margins) for binding by default, uncomment to switch to one side
oneside,
% ngerman for German
english,
% Single line spacing, alternatives: singlespacing, onehalfspacing or doublespacing
onehalfspacing,
% Uncomment to enable draft mode (no pictures, no links, overfull hboxes indicated)
%draft,
% If the document is nolistspacing, onehalfspacing or doublespacing, uncomment this to set spacing in lists to single
onehalfspacing,
% Uncomment to add the list of figures/tables/etc to the table of contents
%liststotoc,
% Uncomment to add the main table of contents to the table of contents
%toctotoc,
% Uncomment to add space between paragraphs
parskip,
% Uncomment to not load the hyperref package
%nohyperref,
% Uncomment to get a line under the header
headsepline,
% Uncomment to place the chapter title next to the number on one line
%chapterinoneline,
% Uncomment to change the layout of the declaration, abstract and acknowledgements pages to match the default layout
%consistentlayout,
% The class file specifying the document structure
]{MastersDoctoralThesis}

% Required for inputting international characters
\usepackage[utf8]{inputenc}
\DeclareUnicodeCharacter{0301}{} % TODO: remove unicode character in .bib

% Output font encoding for international characters
\usepackage[T1]{fontenc}
\usepackage{textcomp} % support the degree symbol

% Use the Palatino font by default
\usepackage{mathpazo}

% Use tu justify paragraphs
\usepackage{ragged2e}

% Import references handler
\usepackage[
    style     = numeric,      % references format style i.e. ieee
    backend   = biber,        % choose the processing program
    natbib    = true,         % enable additional reference formats
    citestyle = numeric-comp, % enable multiple citations
    sortcites = true,         % sort references in multiple citations
    sorting   = none          % sort the reference table
]{biblatex}

% The filename of the bibliography
\addbibresource{mainBiblio.bib}

% Required to generate language-dependent quotes in the bibliography
\usepackage[autostyle=true]{csquotes}

% to set default colours
\usepackage{xcolor}
\definecolor{mygray}{gray}{0.5}

% for gantt charts
\usepackage{tikz}
\usepackage{gantt}

% to rotate pages in landscape
\usepackage{lscape}

% equation arrows with text
\usepackage{amsmath}

% insert degree celsius symbol
\usepackage{textcomp}

% include other pdf pages
\usepackage{pdfpages}

% stop all hyphenation
%\raggedright
\sloppy
\usepackage[none]{hyphenat}

% table out of margin: any suggestion?
\usepackage{tabularx}
\renewcommand\arraystretch{1.5} % Add extra space between table rows

\usepackage{xltabular}
\usepackage{longtable}

% keeps floats ‘in their place’, preventing them from floating past a
% \FloatBarrier command into another section
\usepackage{placeins}

% import code snippets
\usepackage{listings}
\usepackage{xcolor}
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.99,0.99,0.96}
\lstdefinestyle{mystyle}{
  backgroundcolor   = \color{backcolour},
  commentstyle      = \color{codegreen},
  keywordstyle      = \color{magenta},
  numberstyle       = \tiny\color{codegray},
  stringstyle       = \color{codepurple},
  basicstyle        = \ttfamily\small,
  breakatwhitespace = false,
  breaklines        = true,
  captionpos        = b,
  keepspaces        = true,
  numbers           = left,
  numbersep         = 2pt,
  showspaces        = false,
  showstringspaces  = false,
  showtabs          = false,
  tabsize           = 2,
  aboveskip         = 0.1em,
  belowskip         = 0.1em
}
\lstset{style=mystyle}

%----------------------------------------------------------------------------------------
%	MARGIN SETTINGS
%----------------------------------------------------------------------------------------

\geometry{
	paper=a4paper,      % Change to letterpaper for US letter
	inner=2.5cm,        % Inner margin
	outer=3.8cm,        % Outer margin
	bindingoffset=.5cm, % Binding offset
	top=1.5cm,          % Top margin
	bottom=1.5cm,       % Bottom margin
	%showframe,         % Uncomment to show how the type block is set on the page
}

%----------------------------------------------------------------------------------------
%	THESIS INFORMATION
%----------------------------------------------------------------------------------------

% Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
\thesistitle{Search for high-molecular weight linear polymers for the formulation of novel solutions for the fabrication of micro fibers by electromechanical spinning}
% Fabrication of Suspended Nanowires Through Mechano-Near-Field Electrospinning of Polymers in Solution for the Production of Glass-like Carbon

% Your supervisor's name, this is used in the title page, print it elsewhere with \supname and \cosupname
\supervisor{Dr. Héctor Alán Aguirre Soto}
\cosupervisor{Dra. Dora Iliana Medina Medina}

% Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
\examiner{}

% Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
\degree{Master of Science in Nanotechnology (MNT)}

% Your name, this is used in the title page and abstract, print it elsewhere with \authorname
\author{Antonio Osamu Katagiri Tanaka}

% Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname
\addresses{Estado de México, Atizapan de Zaragoza}

% Your subject area, this is not currently used anywhere in the template, print it elsewhere with \subjectname
\subject{Nanotechnology}

% Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
\keywords{nanotechnology, carbon, nano-wires, Near-Field Electrospinning, NFES}

% Your university's name and URL, this is used in the title page and abstract, print it elsewhere with \univname
\university{\href{https://tec.mx/}{Instituto Tecnonólogico y de Estudios Superiores de Monterrey}}

% Your university's campus name and URL, this is used in the title page and abstract, print it elsewhere with \campusname
\campus{\href{https://tec.mx/}{Campus Estado de México}}

% Your department's name and URL, this is used in the title page and abstract, print it elsewhere with \deptname
\department{\href{https://tec.mx/}{School of Engineering and Sciences}}

% Your research group's name and URL, this is used in the title page, print it elsewhere with \groupname
\group{\href{https://www.medinadora.com/people}{ITESM \campusname}}

\examDateAndAddress{\addressname \text{, } \large December 02, 2020}

% Your faculty's name and URL, this is used in the title page and abstract, print it elsewhere with \facname
\faculty{\href{https://samp.itesm.mx/Programas/VistaPrograma?clave=MNT16&modoVista=Default&idioma=ES&cols=0}{Faculty: Nanotechnology}}

\AtBeginDocument{
\hypersetup{pdftitle=\ttitle}          % Set the PDF's title to your title
\hypersetup{pdfauthor=\authorname}     % Set the PDF's author to your name
\hypersetup{pdfkeywords=\keywordnames} % Set the PDF's keywords to your keywords
}

\begin{document}

% Use roman page numbering style (i, ii, iii, iv...) for the pre-content pages
\frontmatter

% Default to the plain heading style until the thesis style is called for the body content
\pagestyle{plain}

%----------------------------------------------------------------------------------------
%	TITLE/COVER PAGE
%----------------------------------------------------------------------------------------

\begin{titlepage}
\begin{center}

\vspace*{.06\textheight}
{\scshape\LARGE \univname\par} % University name

% Research group name and department name
\campusname\\\deptname\\[1cm]

\begin{center}
\includegraphics[width=0.20\textwidth]{./Figures/uniLogo.png}
\end{center}

\vspace{0.5cm}

\HRule \\%[0.5cm]                           % Horizontal line
{\huge \bfseries \ttitle\par}\vspace{0.0cm} % Thesis title
\HRule \\[0.5cm]                            % Horizontal line

\emph{A thesis presented by:}\\

% Author name - remove the \href bracket to remove the link
\href{https://linkedin.com/in/osamu-katagiri-84b2b940/}{\authorname}
%\authorname
\\[0.5cm]

\vfill

% University requirement text
\large \textit{Submitted to the\\
\deptname\\
in partial fulfillment of the requirements for the degree of}\\[0.25cm]
\textit{Master of Science}\\[0.25cm]
\textit{in}\\[0.25cm]
\textit{Nanotechnology}

\vfill

% Address and Date
\examDate\\[4cm]
% University/department logo - uncomment to place it
%\includegraphics{./Figures/uniLogo.png}

\vfill
\end{center}
\end{titlepage}

%----------------------------------------------------------------------------------------
%	COMMITTEE'S PAGE
%----------------------------------------------------------------------------------------

%\includepdf[page=-]{./pdf/AlanDoraSigns}

\begin{committee}
%\addchaptertocentry{\supervisingcommittee} % Add the the supervising committee to the table of contents
The committee members, hereby, certify that have read the dissertation presented by \authorname \text{ } and that it is fully adequate in scope and quality as a partial requirement for the degree of \degreename.

\begin{flushright} \large

\bigskip
\bigskip
\medskip

\noindent \rule[0.0em]{15em}{0.5pt}\\ % This prints a line for the signature
%\href{http://www.alanaguirre.com}{\supname}
\supname\\
Tecnológico de Monterrey\\
\deptname\\
\emph{Principal Advisor}

\bigskip
\bigskip
\medskip

\noindent \rule[0.0em]{15em}{0.5pt}\\ % This prints a line for the signature
%\href{http://www.alanaguirre.com}{\supname}
\cosupname\\
Tecnológico de Monterrey\\
\emph{Co-Advisor and Director of Program}

\bigskip
\bigskip
\medskip

\noindent \rule[0.0em]{15em}{0.5pt}\\ % This prints a line for the signature
%\href{http://www.alanaguirre.com}{\supname}
Dr. Marc Madou\\
University of California, Irvine\\
\emph{Committee Member}

\bigskip
\bigskip
\medskip

\noindent \rule[0.0em]{15em}{0.5pt}\\ % This prints a line for the signature
%\href{http://www.alanaguirre.com}{\supname}
Dr. Jaime Bonilla Ríos\\
Tecnológico de Monterrey\\
\emph{Committee Member}

\end{flushright}

\vfill

\begin{center}

\bigskip
\bigskip
\medskip

\noindent \rule[0.0em]{15em}{0.5pt}\\ % This prints a line for the signature
Dr. Rubén Morales Menéndez \\
\emph{Associate Dean of Graduate Students}\\
\deptname\\
%\href{http://www.doramedina.com}{\supname}

\bigskip
\examDate\\[0cm]

\end{center}
\end{committee}

\cleardoublepage

%----------------------------------------------------------------------------------------
%	DECLARATION PAGE
%----------------------------------------------------------------------------------------

\begin{declaration}
%\addchaptertocentry{\authorshipname} % Add the declaration to the table of contents
\noindent I, \authorname, declare that this thesis titled, \enquote{\ttitle} and the work presented in it are my own. I confirm that:

\begin{itemize} 
\item This work was done wholly or mainly while in candidature for a research degree at this University.
\item	Where any part of this dissertation has previously been submitted for a degree or any other qualification at this University or any other institution, this has been clearly stated.
\item	Where I have consulted the published work of others, this is always clearly attributed.
\item	Where I have quoted from the work of others, the source is always given. With the exception of such quotations, this dissertation is entirely my own work.
\item	I have acknowledged all main sources of help.
\item	Where the dissertation is based on work done by myself jointly with others, I have made clear exactly what was done by others and what I have contributed myself.
\\
\end{itemize}

\bigskip

\begin{flushright}
\noindent \rule[0.0em]{15em}{0.5pt}\\ % This prints a line for the signature
\authorname\\
\examDate
\end{flushright}

\vfill

\begin{center}
@2020 by \authorname\\
All rights reserved
\end{center}

\end{declaration}

\cleardoublepage

%----------------------------------------------------------------------------------------
%	QUOTATION PAGE
%----------------------------------------------------------------------------------------

\vspace*{\fill}

\begin{center}
%\includegraphics[width=0.60\textwidth]{./Figures/quote.png}

\noindent\enquote{\itshape Carbon is a simple element but ............... \\
\medskip 
one branch of chemistry is devoted to its compounds! \\
\medskip 
... one branch of science is devoted to the many forms of the element as a solid material. \\
\medskip 
The best of this is that although most carbon materials are grey or black to the naked eye and the uninitiated. A closer examination reveals the form, beauty and even color of carbon science.}\bigbreak
\end{center}

\begin{flushright}
\textbf{Marsh, Harry} \linebreak
\emph{Universitat d'Alacant, Alicante, Spain} \linebreak
\href{https://www.scopus.com/authid/detail.uri?authorId=7201585695}{scopus.com}
\end{flushright}

\vspace*{\fill}

%----------------------------------------------------------------------------------------
%	DEDICATION
%----------------------------------------------------------------------------------------

\begin{dedication}
%\addchaptertocentry{\dedicationname} % Add the acknowledgements to the table of contents
To my parents, who believed in me no matter the outcome. Mom, dad, this work is possible because of you, for always telling me to believe in my dreams.

Thanks for all your unconditional confidence, support, patience, and encouragement. You were my main motivation for pushing through this work.

\end{dedication}

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\begin{acknowledgements}
%\addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
There are a number of people without whom this thesis might not have been written,
and to whom I am greatly indebted.\\

% To my mother, Helena, who continues to learn, grow and develop and who has been a source of encouragement and inspiration to me throughout my life, a very special thank you for providing a ‘writing space’ and for nurturing me through the months of writing. And also for the myriad of ways in which, throughout my life, you have actively supported me in my determination to find and realise my potential, and to make this contribution to our world.\\

% To my mother-in-law, Maryam, known only briefly but loved and missed, who represented to me ‘living proof’ of Black women’s ability to redefine and recreate our lives despite, and maybe even because of, the tremendously constraining, oppressive and repressive situations in which we often exist.\\

% To my dear husband, Bala who remains willing to engage with the struggle, and ensuing discomfort, of having a partner who refuses to accept the given role of the "Black woman" and is actively engaged in redefining and redesigning that role. A very special thank you for your practical and emotional support as I added the roles of wife and then mother, to the competing demands of business, work, study and personal development.\\

% Thanks to dear Kamal and Azaria, for being so supportive - even when being ‘without Mum’ was hard, and for your help with the bibliography and diagrams. This work is for, and because of you and all the generations to come. It is dedicated to all our journeys in learning to thrive.\\

% It is also dedicated to Josephine - friend, ‘sister’, colleague, ‘co- traveller’ and researcher - who knowingly and unknowingly- led me to an understanding of some of the more subtle challenges to our ability to thrive. If our attempts to claim our right to speak our truth, and to unravel and follow the threads through which our oppression is maintained, and are instrumental in helping one other Black woman from ‘going over the edge’ - in which sight we continuously live our lives - perhaps, it might be seen that your invaluable contribution to the attainment of many of the insights gained was worth it. It is also written in honour of all our my grandmothers, and their mothers, and grandmothers, and to all our and mothers and grandmothers, who though not marked by history struggled so heroically to resist the definitions attributed to them by the dominant system and communicate messages of hope and expectation. I am also very grateful to Maya Angelou, for the inspiration she has been to methrough her writings, poetry readings and as a living representation of a woman thriving. Thanks to Mary Washington, Alice Walker, Jackie Kay, Buchi Emecheta, Toni Cade, Lauretta Ngobo, bell hooks, and other Black women writers of the 1980’s, who in telling stories for and about us validated me and my experiences in ways I cannot begin to explain here.\\

% Loving thanks to my friends and learning partners, Agnes, Susan, Ulana and Namonya, who played such important roles along the journey, as we mutually engaged in making sense of the various challenges we faced and in providing encouragement to each other at those times when it seemed impossible to continue.

I offer my gratitude and appreciation to my supervisors, \supname \textit{ } and \cosupname, for the deft ways in which you fondly challenged and supported me through out the whole of this work - knowing when to push and when to let up. I offer special thanks to those who supported me in the mechanics of producing this thesis. \supname, for reading and rereading drafts, editing and proofing, typing and helping me get ‘unstuck’ with this thesis on many occasions. And Braulio and Arnoldo for ‘rescuing’ me at those times when I was almost defeated by the laboratory equipment.

I would also like to acknowledge the support from Tecnológico de Monterrey and CONACyT (tuiton and living expenses). I also thank Regina Elizabeth Vargas from CIDyT, for the countless hours she helped me with sample characterization.

Finally, I would like to thank my family, who always believed in me. I am grateful for their love, caring and patience even in the hardest and restless days. They always encourage me to following my dreams, and that feeling will always be dear to me.

\end{acknowledgements}

%----------------------------------------------------------------------------------------
%	ABSTRACT PAGE
%----------------------------------------------------------------------------------------

\begin{abstract}
\addchaptertocentry{\abstractname} % Add the abstract to the table of contents

Carbon nano-wires are versatile structures composed of carbon chains with a wide range of applications due to their high chemical resistance and electric properties. Regardless of the high interest in the implementation of carbon nano-wires in energy, environmental and health-care applications, no feasible processes have been developed to fabricate carbon nano-wires with spatial control at a reasonable cost. Carbon nano-wires have been fabricated with the use of a photoresist, but little is known about polymers that can produce conductive carbon nano-wires after pyrolysis. Various polymer solutions have been tested in near field electrospinning (NFES) and photopolymerization separately, however, few have been tested for nano-wire fabrication purposes through the process of spatio-temporal deposition with NFES, photo-polymerization for cross-linking and pyrolysis. The intention behind the thesis proposal is to use rheological analyses of different polymer solutions to determine if they can be easily electrospun at low voltages and then fabricate nano-wires with them. This thesis work arises from the need to test a greater variety of polymer-solvent combinations with the goal of designing a polymer solution to fabricate carbon nano-wires with higher conductivity than the current SU-8 polymeric nano-fibers. The present work includes the design of polymer solutions that can be electrospun, with the hope that the selected high molecular weight polymers can be photo-polymerized, and then pyrolyzed into conducting carbon nanowires. The overarching goal is to contribute towards the development of novel designed polymer solutions to achieve mass scale manufacturing of conductive carbon nano-wires in an inexpensive, continuous, simple and reproducible manner as central components for nano-sensors. \\

\noindent
\textbf{keywords:} \keywordnames
\end{abstract}

%----------------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES PAGES
%----------------------------------------------------------------------------------------

\listoffigures % Prints the list of figures
\addchaptertocentry{\listfigurename} % Add to the table of contents

\listoftables % Prints the list of tables
\addchaptertocentry{\listtablename} % Add to the table of contents

\tableofcontents % Prints the main table of contents

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - CHAPTERS
%----------------------------------------------------------------------------------------

\mainmatter % Begin numeric (1,2,3...) page numbering

\pagestyle{thesis} % Return the page headers back to the "thesis" style

% Include the chapters of the thesis as separate files from the Chapters folder
% Uncomment the lines as you write the chapters

\setlength{\parskip}{1.5em} % Paragraph spacing
%\RaggedRight
{\fontsize{12}{15} \selectfont

\include{./Chapters/1}
\include{./Chapters/2}
\include{./Chapters/3}
\include{./Chapters/4}
\include{./Chapters/5}

}

%\include{Chapters/__IntroToLaTeX}

\appendix % Cue to tell LaTeX that the following "chapters" are Appendices

% Include the appendices of the thesis as separate files from the Appendices folder
% Uncomment the lines as you write the Appendices

%----------------------------------------------------------------------------------------
%	ABBREVIATIONS
%----------------------------------------------------------------------------------------

\begin{abbreviations}{ll} % Include a list of abbreviations (a table of two columns)

\textbf{CEM}   & \textbf{C}ampus \textbf{E}stado de \textbf{M}éxico\\
\textbf{CNWs}  & \textbf{C}arbon \textbf{N}ano-\textbf{w}ire\textbf{s}\\
\textbf{DC}    & \textbf{D}irect \textbf{C}urrent\\
\textbf{EMS}   & \textbf{E}lectro\textbf{m}echanical \textbf{S}pinning\\
\textbf{FFES}  & \textbf{F}ar \textbf{F}ield de \textbf{E}lectro\textbf{s}pinning\\
\textbf{ITESM} & \textbf{I}nstituto \textbf{T}ecnonólogico y de \textbf{E}studios \textbf{S}uperiores de \textbf{M}onterrey\\
\textbf{MA}    & \textbf{Ma}ssachusetts\\
\textbf{MEMS}  & \textbf{M}icro\textbf{e}lectro\textbf{m}echanical \textbf{S}ystems\\
\textbf{MNT}   & \textbf{M}aestría en \textbf{N}ano\textbf{t}ecnología \emph{(Master of Science in Nanotechnology)}\\
\textbf{MTY}   & \textbf{M}on\textbf{t}erre\textbf{y} \emph{or} Campus \textbf{M}on\textbf{t}erre\textbf{y}\\
\textbf{NFEMS} & \textbf{N}ear-\textbf{F}ield \textbf{E}lectro\textbf{m}echanical \textbf{S}pinning\\
\textbf{NFES}  & \textbf{N}ear \textbf{F}ield de \textbf{E}lectro\textbf{s}pinning\\
\textbf{USA}   & \textbf{U}nited \textbf{S}tates of \textbf{A}merica\\
\textbf{UV}    & \textbf{U}ltra\textbf{v}iolet\\

\end{abbreviations}

%----------------------------------------------------------------------------------------
%	SYMBOLS
%----------------------------------------------------------------------------------------

\begin{symbols}{lll} % Include a list of Symbols (a three column table)

%$a$ & distance & \si{\meter} \\
%$P$ & power & \si{\watt} (\si{\joule\per\second}) \\
Symbol   & Name              & Unit \\
\addlinespace % Gap to separate the Roman symbols from the Greek
$\textrm{Mw}$  & Molecular Weight             & $\textrm{g} / \textrm{mole}$ \\
$\eta_0$       & Zero-shear Viscosity         & $\textrm{Pa} \cdot \textrm{s}$	\\
$K$            & Electrical Conductivity      & $\textrm{S} / \textrm{m}$	\\
$\gamma$       & Surface Tension              & $\textrm{N} / \textrm{m}$ \\
$\dot{\gamma}$ & Shear Rate                   & $1 / \textrm{s}$ \\
$\rho$         & Density                      & $\textrm{Kg} / \textrm{m}^3$ \\
$C_{polymer}$  & Polymer concentration        & $\textrm{wt\%}$	\\
$D_{nozzle}$   & Nozzle Inner Diameter        & $\textrm{m}$ \\
$Q$            & Flow Rate                    & $\textrm{m}^3 / \textrm{s}$	\\
$L$            & NFES Working Distance        & $\textrm{m}$	\\
$\Phi_0$       & NFES Applied Voltage         & $\textrm{V}$	\\
$v_{stage}$    & Collector/Stage velocity     & $\textrm{m} / \textrm{s}$ \\
$D_{fiber}$    & Fiber Diameter               & $\textrm{m}$ \\
$FiberGap$     & Distance between fibers      & $\textrm{m}$ \\
$R_{jet}$      & NFES Jet Radius              & $\textrm{m}$ \\
$Oh$           & Ohnesorge number             & NA. \\
$Pe$           & Peclet number                & NA. \\
$Re$           & Reynolds number              & NA. \\
$We$           & Weber number                 & NA. \\
$\Psi$         & Dimensionless Field Strength & NA. \\

\end{symbols}

%----------------------------------------------------------------------------------------
%	PHYSICAL CONSTANTS/OTHER DEFINITIONS
%----------------------------------------------------------------------------------------

%\begin{constants}{lr@{${}={}$}l} % The list of physical constants is a three column table

% The \SI{}{} command is provided by the siunitx package, see its documentation for instructions on how to use it

%Speed of Light & $c_{0}$ & \SI{2.99792458e8}{\meter\per\second} (exact)\\
%Constant Name & $Symbol$ & $Constant Value$ with units\\

%\end{constants}

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - APPENDICES
%----------------------------------------------------------------------------------------

\include{Appendices/Appendix_ViscosityCurves}
\include{Appendices/Appendix_CriticalConcentrations}
\include{Appendices/Appendix_MicroscopyChar}
\include{Appendices/Appendix_FiberDiameters}
\include{Appendices/Appendix_ReviewTable}
\include{Appendices/Appendix_ImageAnalysisCode}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\printbibliography[heading=bibintoc, title={Bibliography}]
% \addcontentsline{toc}{chapter}{Bibliography}

%% https://www.overleaf.com/learn/latex/bibtex_bibliography_styles
%% abbrv, acm, alpha, apalike, ieeetr, plain, siam, unsrt
%\bibliographystyle{plain}
%\renewcommand\bibname{References}
%\bibliography{mainBiblio}

%----------------------------------------------------------------------------------------
%	CURRICULUM VITAE
%----------------------------------------------------------------------------------------

% \includepdf[page=-,fitpaper=false,landscape=false]{./pdf/CV/CV}

%----------------------------------------------------------------------------------------

\end{document}